### 2022-09-03 Writing/figures.Rmd
initial commit

<!-- Geographic focus -->

```{r results = "hide"}
#=== HPA boundary ===#
hpa <- st_read(here("Data/RawData/hp_bound2010.shp")) %>% 
  st_simplify()

us_states_borders <- states(cb = TRUE) %>%
  st_as_sf() %>%
  st_transform(st_crs(hpa)) %>% 
  rename(state_code = STATEFP, state_abb = STUSPS) %>%
  dplyr::select(state_code, state_abb)

```

```{r gfocus, fig.cap = "Geographic Focus of the Study by Crop Type", fig.height = 4, fig.width = 6.5}

all_counties <- readRDS(here("Data/ProcessedData/base_counties.rds"))

# results$states

gfocus_data <- results %>% 
  dplyr::select(crop_type, states) %>% 
  mutate(plot_data = list(
    all_counties[state %in% states, ] 
  )) %>% 
  dplyr::select(-states) %>% 
  unnest() %>% 
  dplyr::select(crop_type, geometry, state_name) %>% 
  st_as_sf()

gfocus_state_borders <- results %>% 
  dplyr::select(crop_type, states) %>% 
  mutate(state_data = list(
    filter(us_states_borders, state_abb %in% states)
  )) %>% 
  dplyr::select(-states) %>% 
  unnest() %>% 
  st_as_sf()

t_gfocus <- tm_shape(hpa) +
  tm_polygons(col = "blue", alpha = 0.3) +
tm_shape(gfocus_data) +
  tm_borders() +
tm_shape(gfocus_state_borders) +
  tm_borders(lwd = 1.5, col = "red") +
  tm_facets(by = "crop_type", nrow = 1)

# tmap_save(
#   t_gfocus,
#   here("Writing/Figures", "geogprahic_focus.pdf"),
#   height = 3.5,
#   width = 6
# )

t_gfocus

```

`r run_pagebreak()`

<!-- The impact of saturated thickness on yield (Intensive Margin) -->

```{r yield-intensive, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}

g_intensive$data_for_plot[[1]] %>% 
  .[days_ab_35 == median(days_ab_35), ]

g_intensive$data_for_plot[[1]] %>% 
  .[days_ab_35 == median(days_ab_35), ]

g_intensive$data_for_plot[[1]]

str(g_intensive$data_for_plot[[1]])

g_intensive <- reg_results %>% 
  dplyr::select(crop_type, data_for_plot) %>% 
  mutate(data_for_plot_balance = list(
    data_for_plot[days_ab_35 == median(days_ab_35), ]
  )) 
  unnest(cols = "data_for_plot") %>% 
  ggplot(data = .) +
    geom_ribbon(
      aes(
        ymin = yhat_d, 
        ymax = yhat_u, 
        x = balance
      ),
      alpha = 0.4
    ) +
    geom_line(aes(y = y_hat, x = balance)) +
    facet_grid(crop_type ~ sat, scales = "free_y") +
    theme_bw()

# ggsave(
#   here("Writing/Figures", "yield_intensive.pdf"),
#   g_intensive, 
#   height = 3.5,
#   width = 6
# )

g_intensive

```

`r run_pagebreak()`

<!-- The impact of saturated thickness on Irrigation Share -->

```{r ir-share, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}

g_ir_share <- results_all %>% 
  dplyr::select(crop_type, ir_share_eval_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = ir_share_hat, 
      x = sat, 
      color = factor(balance_avg)
    )
  ) +
  facet_grid(. ~ crop_type) +
  theme_bw()

# ggsave(
#   here("Writing/Figures", "ir_share.pdf"),
#   g_ir_share, 
#   height = 3.5,
#   width = 6
# )

g_ir_share
```

`r run_pagebreak()`

<!-- The impact of saturated thickness on yield (Combined) -->

```{r yield-combined, fig.cap = "The Impact of Saturated Thickness on Yield", fig.height = 6, fig.width = 6.5}

g_sat_impact <- results_all %>% 
  dplyr::select(crop_type, sat_sim_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = mean_yield, 
      x = balance, 
      color = factor(sat)
    )
  ) +
  facet_grid(crop_type ~ ., scales = "free_y") +
  scale_color_viridis_d(name = "Saturated Thickness") +
  theme_bw() +
  theme(
    legend.position = "bottom"
  )

# ggsave(
#   here("Writing/Figures", "yield_combined.pdf"),
#   g_sat_impact, 
#   height = 3.5,
#   width = 6
# )

g_sat_impact
```

```{r }
data_sim_yield <- readRDS(here("Results/simulated_yield_data.rds"))

cl_data_current <- readRDS(here("Data/ProcessedData", "summrized_gridMET.rds"))

cl_data_future <- 
  expand.grid(
    model = model_ls,
    rcp = rcp_ls
  ) %>% 
  data.table() %>% 
  rowwise() %>% 
  #=== define maca data file names ===#
  mutate(file_name = list(
    paste0(
      "Data/ProcessedData/MACA/sum_", 
      model,
      "_",
      rcp,
      "_county.rds"
    )  
  )) %>% 
  mutate(maca_data = list(
    #=== read maca data ===#
    readRDS(file_name) %>% 
    #=== calculate 10-year balance avg ===#
    .[, 
      `:=`(
        balance_avg = mean(balance), 
        days_ab_35_avg = mean(days_ab_35)
      ), 
      by = .(floor((year - 2020) / 10), sc_code)
    ] 
  )) %>% 
  dplyr::select(model, rcp, maca_data) %>% 
  unnest() %>% 
  data.table()

cl_data <- 
  rbind(
    cl_data_current[, .(sc_code, year, balance, days_ab_35, model, rcp)],
    cl_data_future[, .(sc_code, year, balance, days_ab_35, model, rcp)]
  ) %>% 
  .[order(sc_code, year), ]

ggplot(data = cl_data_future) +
  geom_boxplot(aes(y = balance, x = factor(year))) +
  facet_grid(model ~ rcp)

felm(balance ~ year , data = cl_data_future) %>% summary()

```