### 2023-07-18 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewer_1.rmd
revision started

---
output:
  pdf_document:
    keep_tex: true
---
Reviewer #1 (Comments for the Author):

This manuscript assesses the impact of aquifer thickness on agricultural production during drought in the area overlying the High Plains aquifer in the central United States. The authors use crop, meteorological, and aquifer thickness data in regressions to develop relationships among these quantities. Their most important finding is the impact of aquifer thickness on productivity during drought and the ramifications for aquifer conservation efforts. I have read this manuscript over a number of times and found it of considerable interest and worthy of publication in Nature Water. However, I do have a few concerns that I wanted to bring to the attention of the authors. My most important concerns are as follows (not necessarily in order of importance):

\textcolor{blue}{We really appreciate your comments. Addressing them has made our paper significantly stronger. Please find our replies to your comments written in blue.}

1. Aquifer thickness versus saturated thickness – The authors speak of "aquifer saturated thickness", which is a redundant term. Remember, the definition of an aquifer is a saturated unit – there is no unsaturated aquifer. Please use "aquifer thickness" to avoid propagating this confusing usage into the future.

\textcolor{blue}{All the instances of "saturated thickness" and "aquifer saturated thickness" have been replaced by "aquifer thickness".}

2. Dependence on aquifer thickness – I had a number of questions about the relationships displayed in Figure 1. First, the regression appears to have been done year by year. Are the quantiles recalculated each year to reflect that there may be changes between categories as time goes on? Second, when the authors state the lowest quantile goes from 0 to 33% are they assuming that the minimum thickness of 9 m is the zero point? Third, I assume that the county and year fixed effects are calculated as part of the regression. Are there any insights that can be gained from the interpretation of those quantities? Fourth, the authors need to point out that the same aquifer thickness can result in dramatically different transmissive characteristics. Thus, a relatively thin aquifer interval composed of coarse sands and gravels may be more resilient to drought than a much thicker aquifer dominated by fine sands and silts. Could that be a major player in the error bars in the plots? Please mention this at some point in the manuscript.

- First point: \textcolor{blue}{Quantiles are not recalculated each each. If we do so, then the same quantile would represent different aquifer thickness range every year. When modeling the impact of aquifer thickness, this is not desirable because we are implicitly making assumptions that different ranges of aquifer thickness would have the same impact.}
-  Second point: \textcolor{blue}{Actually, 0% quntile is 12 m. "9m" was repalced with "12m" in the current manuscript.}
-  Third point: \textcolor{blue}{No interesting insights will be gained from them because they are merely shifters without any intractions with aquifer thickness. They tell us how much higher yield in a single county relative to another county on average. But, we do not beleive that is useful information at least in this artcile.}
-  Fourth point: \textcolor{blue}{This is a very important point. We have added texts describing the implications of this fact in page xx.}

3. Conditioned on drought risk exposure – I wasn’t sure what the authors meant by this in the second paragraph of Section 2.2. This must be related to the water deficit but the lack of explanation in the main text or the methods section left me wondering. Please clarify at some point in the manuscript.

\textcolor{blue}{What we meant by "conditiond on drought risk exposure" is "after taking into account the impact of climate variables." In order to avoid confusion, we simply removed it from the sentence in the current manuscript.}

4. Equation 2 – Please define all terms in the equation and provide an explanation of what is meant by "estimated semi-parametrically".

\textcolor{blue}{We now have a more detailed mathematical expressions of the model with all the terms explained. We eliminated the word semi-parametrically as it is not ncecessary. Generalized additive model that estiamtes the impact of a variable using splines without functional form assumption are sometimes referred to semi-parametric regression. That's what we meant by "estimate semi-parametrically" in our previous manuscript.}

5. Kansas and California efforts – I’ve seen a number of papers describing efforts in Kansas and California to address aquifer depletion via LEMAs (Kansas) and SGMA (California). The 1985-2016 period examined by the authors was before most of the efforts in Kansas got started. My understanding is that the first LEMA was established in 2013 but then no more were set up until after 2016. This manuscript would be more impactful if the authors could speculate on how their relationships might change with greater attention to groundwater conservation. For example, I would think the decrease in yield the authors observe with negative water deficits would disappear when the irrigators are using soil moisture sensors and are focused on water-use efficiency. I would expect changes as well for tipping points.

\textcolor{blue}{Tim?}

## Minor points

+ Section 1, 3rd Paragraph – "Two aquifer pathways exist through aquifer depletion will.." – confusing sentence – please replace "through" with "by which".

\textcolor{blue}{Corrected.}

+ Section 1, 4th Paragraph – "we generate empirical evidence" seems awkward to me. How about replacing with "we present empirical evidence"? Next sentence – "Our analysis advances on previous…" reads better as "Our analysis extends previous…" There are many issues such as these and the previous comment. Please address such issues so that the manuscript is an easier read.

\textcolor{blue}{We completely agree. Both of the suggested changes have been made.}

+ Section 2.1 title – I would delete "marginally" as it is not needed and some will find it confusing.

\textcolor{blue}{Tim, any thoughts on this. I feel like it is still important to say the increase is not much.}

+ Figure 2 - Why do the plots stop at a thickness of about 12 m when you had earlier defined the minimum thickness as 9 m and mention 10 m in the text when discussing the relationships?

\textcolor{blue}{Thank you for catching this. 9m and 10m should have been 12m.}

+ Section 2.3, First Paragraph – please clarify that "average per-area production" is a combination of irrigated and rainfed yields. That would make the manuscript easier to follow.

\textcolor{blue}{We added "(weighted-average of irrigated and rainfed yields, where weights are production area shares)" so that it is clear what average per-area production means.}

+ Figure 5 – I assume that the left plot is corn and the right soybean, but please clarify.

\textcolor{blue}{Thanks for noticing this. Fixed.}

+ Data and Code Availability – Are the data from Dr. Haacker on the Github repository? Please place them there prior to publication so future investigators can access them rather than having to track down Dr. Haacker. The data set is over seven years old so it should be made available according to the principles of open science.

\textcolor{blue}{Yes, we will make the data available online.}

+ Line before Equation (2) – "REF" should be replaced with the actual reference.

\textcolor{blue}{Tim?}

+ Discussion, First Paragraph – "reduction in aquifer depletion" should be "reduction in aquifer thickness".

\textcolor{blue}{Thank you for catching this. Fixed.}

+ Section 4.2, 3rd Paragraph – "the the impact" – remember to spell and grammar check prior to submission.

\textcolor{blue}{Thank you for catching this. Fixed.}

+ Reference in Section 3 – I doubt that "Butler Jr" is the surname, please replace with "Butler". Suffixes apparently aren’t handled well by some of the bibliography generation programs.

\textcolor{blue}{}

+ Please remember to close all parentheses and don't mispair brackets and parentheses. Such stuff can raise questions about the care used in the analysis.