### 2023-07-18 Writing/NatureWater/1stRoundRevision/revised_manuscript/cover_letter.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather},
  pdfauthor={Taro Mieno; Timothy Foster; Shunkei Kakimoto; Nicholas Brozovic},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Aquifer depletion exacerbates agricultural drought losses in the US High Plains}
\author{Taro Mieno\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:tmieno2@unl.edu}{\nolinkurl{tmieno2@unl.edu}}} \and Timothy Foster\footnote{Department of Mechanical, Aerospace and Civil Engineering, University of Manchester, \href{mailto:timothy.foster@manchester.ac.uk}{\nolinkurl{timothy.foster@manchester.ac.uk}}} \and Shunkei Kakimoto\footnote{Department of Applied Economics, University of Minnesota, \href{mailto:kakim002@umn.edu}{\nolinkurl{kakim002@umn.edu}}} \and Nicholas Brozovic\footnote{Department of Agricultural Economics, University of Nebraska Lincoln, \href{mailto:nbrozovic@nebraska.edu}{\nolinkurl{nbrozovic@nebraska.edu}}}}
\date{}

\begin{document}
\maketitle

Dear Editor,\\

Thank you for considering this manuscript for publication at Nature Water. In this study, we demonstrate the importance of aquifer conservation for ensuring agricultural resilience to current and future drought events. Using the High Plains Aquifer in the central United States as a case study, we show that declines in aquifer saturated thickness significantly reduces the overall productivity of both corn and soybean during drought events. Negative impacts on agricultural productivity during droughts occur both through reductions in per-area crop yields, and reductions in expected irrigated areas as farmers are forced to transition land to rainfed production as a result of water supply constraints caused by aquifer depletion. Our findings demonstrate that unsustainable groundwater use is likely to substantially increase vulnerability of agriculture to drought in many parts of the world, and highlight the need for greater focus on proactive conservation of groundwater stocks to support adaptation of farms, agricultural supply chains, and rural economies to climate change.\\

Our work is significant because we quantify the impact of a decline in saturated thickness of groundwater on the resiliency of agricultural production, which has rarely been quantified in the literature before. We confirm quantitatively the importance of keeping relatively high saturated thickness for sustainable high-productivity agricultural production. While our results are for the US High Plains, the insights we generated here are likely to hold elsewhere.\\

This article does not have any realated manuscripts that are under consideration. However, its pre-print has been published at EarthArXiv (\url{https://eartharxiv.org/repository/view/5126/}). We have not had any prior discussions with a Nature Water editor about this work. \\

Here is the list of potential reviewers that we may recommend.

\begin{itemize}
    \item James Butler (\url{https://geo.ku.edu/people/james-k-butler-jr})
    \item Marc Muller (\url{https://engineering.nd.edu/faculty/marc-francois-muller/})
    \item Sam Zipper (\url{https://www.samzipper.com/authors/samuel-c.-zipper/})
    \item Christian Klassert (\url{https://www.ufz.de/index.php?en=38767})
    \item Jen Burney (\url{http://www.jaburney.net/})
\end{itemize}

We do not have any reviewers that we would like to avoid.

\end{document}