### 2023-07-21 Codes/Retired/3_1_make_figures_results.Rmd

---
title: "Make Figures"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi=400,
  fig.width=7.5,
  fig.height=5,
  out.width="750px",
  out.height="500px"
)
```

# Prep

```{r package, include = F}
library(tmap)
library(tidyverse)
library(data.table)
library(tidycensus)

#=== regression results ===#
reg_results <- readRDS(here("Shared/Results/reg_results.rds"))

#=== hypothetical sim for regression results illustration ===#
sim_results <- readRDS(here("Shared/Results", "hypothetical_sim_results.rds"))

#=== simulation results ===#
data_sim_yield <- readRDS(here("Shared/Results", "simulated_yield_data.rds"))

#=== county sf ===#
data(fips_codes)

all_counties <-
  tigris::counties() %>% 
  rename(
    state_code = STATEFP, 
    county_code = COUNTYFP,
    county_name = NAME
  ) %>% 
  left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
  dplyr::select(state_code, state, county_code, county_name) %>% 
  mutate(sc_code = paste0(state_code, county_code)) %>% 
  data.table()
```

# Yield response to drought and heat
## Water Deficit (Balance) on Yield by Saturated Thickness Level

```{r yield-intensive-balance, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}
(
g_intensive_balance <- 
  sim_results %>% 
  filter(model_name == "balance_d35") %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  mutate(days_ab_35_med = 
    yield_hat_eval_data[which.min(abs(days_ab_35-median(days_ab_35))), days_ab_35]
  ) %>% 
  mutate(yield_hat_eval_data = list(
    yield_hat_eval_data[days_ab_35 == days_ab_35_med, ]
  )) %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  unnest(cols = "yield_hat_eval_data") %>% 
  ggplot(data = .) +
    geom_ribbon(
      aes(
        ymin = yhat_d, 
        ymax = yhat_u, 
        x = balance
      ),
      alpha = 0.4
    ) +
    geom_line(aes(y = y_hat, x = balance)) +
    facet_grid(crop_type ~ sat, scales = "free_y") +
    theme_bw()
)

# (
# g_intensive_balance <- 
#   sim_resulits %>% 
#   filter(model_name == "et0_prcp") %>% 
#   dplyr::select(crop_type, yield_hat_eval_data) %>% 
#   mutate(prcp_med = 
#     yield_hat_eval_data[which.min(abs(prcp-median(prcp))), prcp]
#   ) %>% 
#   mutate(yield_hat_eval_data = list(
#     yield_hat_eval_data[prcp == prcp_med, ]
#   )) %>% 
#   dplyr::select(crop_type, yield_hat_eval_data) %>% 
#   unnest(cols = "yield_hat_eval_data") %>% 
#   ggplot(data = .) +
#     geom_ribbon(
#       aes(
#         ymin = yhat_d, 
#         ymax = yhat_u, 
#         x = et0
#       ),
#       alpha = 0.4
#     ) +
#     geom_line(aes(y = y_hat, x = et0)) +
#     facet_grid(crop_type ~ sat, scales = "free_y") +
#     theme_bw()
# )
```

## Days above 35 on Yield by Saturated Thickness Level

```{r yield-intensive-d35, fig.cap = "The Impact of the Number of Days above 35 on Irrigated Yield at Various Saturated Thickness Levels", fig.height = 4, fig.width = 6.5}
(
g_intensive_d35 <- 
  sim_results %>% 
  filter(model_name == "balance_d35") %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  mutate(balance_med = 
    yield_hat_eval_data[which.min(abs(balance-median(balance))), balance]
  ) %>% 
  mutate(yield_hat_eval_data = list(
    yield_hat_eval_data[balance == balance_med, ] %>% 
      .[, type := "days_ab_35"]
  )) %>% 
  dplyr::select(crop_type, yield_hat_eval_data) %>% 
  unnest(cols = "yield_hat_eval_data") %>% 
  ggplot(data = .) +
    geom_ribbon(
      aes(
        ymin = yhat_d, 
        ymax = yhat_u, 
        x = days_ab_35
      ),
      alpha = 0.4
    ) +
    geom_line(aes(y = y_hat, x = days_ab_35)) +
    facet_grid(crop_type ~ sat, scales = "free_y") +
    theme_bw()
)
```

# The impact of saturated thickness on Irrigation Share

```{r ir-share, fig.cap = "The Impact of Saturated Thickness on Irrigated Yield", fig.height = 4, fig.width = 6.5}

g_ir_share <- 
  sim_results %>% 
  dplyr::select(crop_type, ir_share_eval_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = ir_share_hat, 
      x = sat, 
      color = factor(balance_avg)
    )
  ) +
  facet_grid(. ~ crop_type) +
  theme_bw()

# ggsave(
#   here("Writing/Figures", "ir_share.pdf"),
#   g_ir_share, 
#   height = 3.5,
#   width = 6
# )

g_ir_share
```

# The Impact of Water Deficit on Yield by Saturated thickness (intensive and extensive margins combined)

```{r yield-combined, fig.cap = "The Impact of Saturated Thickness on Yield", fig.height = 6, fig.width = 6.5}

g_sat_impact <- 
  sim_results %>% 
  filter(model_name == "balance_d35") %>% 
  dplyr::select(crop_type, avg_yield_hat_eval_data) %>% 
  mutate(days_ab_35_med = 
    avg_yield_hat_eval_data[which.min(abs(days_ab_35-median(days_ab_35))), days_ab_35]
  ) %>% 
  mutate(balance_avg_med = 
    avg_yield_hat_eval_data[which.min(abs(balance_avg-median(balance_avg))), balance_avg]
  ) %>% 
  mutate(days_ab_35_avg_med = 
    avg_yield_hat_eval_data[which.min(abs(days_ab_35_avg-median(days_ab_35_avg))), days_ab_35_avg]
  ) %>% 
  mutate(plot_data = list(
    avg_yield_hat_eval_data %>% 
    .[days_ab_35 == days_ab_35_med, ] %>% 
    .[balance_avg == balance_avg_med, ] %>% 
    .[days_ab_35_avg == days_ab_35_avg_med, ] 
  )) %>% 
  dplyr::select(crop_type, plot_data) %>% 
  unnest() %>% 
  ggplot(data = .) +
  geom_line(
    aes(
      y = mean_yield, 
      x = balance, 
      color = factor(sat)
    )
  ) +
  facet_grid(crop_type ~ ., scales = "free_y") +
  scale_color_viridis_d(name = "Saturated Thickness") +
  theme_bw() +
  theme(
    legend.position = "bottom"
  )

g_sat_impact
```

# Climate change impact

```{r echo = F, eval = F}
data_sim_yield <- 
  here("Shared/Results/simulated_yield_data.rds") %>% 
  readRDS() %>% 
  data.table() %>% 
  .[, cl_period := case_when(
    real_year <= 2020 ~ "past",
    real_year > 2020 & real_year <= 2040 ~ "2021-2040",
    real_year > 2040 & real_year <= 2060 ~ "2041-2060",
    real_year > 2060 & real_year <= 2080 ~ "2061-2080",
    real_year > 2080 & real_year <= 2100 ~ "2081-2099",
  )] %>% 
  .[, yield := y_hat_dr * (1 - ir_share) + y_hat * ir_share]
```

## Overall trend (Corn)

```{r echo = F, eval = F}
plot_data_corn_ir <- 
  data_sim_yield %>% 
  .[crop_type == "corn", ] %>% 
  .[sat_type == "hypothetical", ] %>% 
  .[, .(yield, y_hat_dr, cl_period, sc_code, rcp, real_year)]

plot_data_corn_dr <-
  copy(plot_data_corn_ir) %>% 
  .[, yield := NULL] %>% 
  setnames("y_hat_dr", "yield") %>% 
  unique(by = c("cl_period", "sc_code", "rcp", "real_year"))

ggplot(plot_data_corn) +
  geom_boxplot(aes(y = yield, x = cl_period, fill = factor(sat))) +
  facet_grid(rcp ~ .)

ggplot(data_overall[crop_type == "corn", ]) +
  geom_boxplot(aes(y = yield, x = cl_period, fill = factor(sat))) +
  facet_grid(rcp ~ .)

data_overall <- 
  data_sim_yield %>% 
  .[sat_type == "hypothetical", ] %>% 
  .[, .(yield = mean(yield)), by = .(rcp, sc_code, cl_period, crop_type, cl_type, sat)]

ggplot(data_overall[crop_type == "corn", ]) +
  geom_boxplot(aes(y = yield, x = cl_period, fill = factor(sat))) +
  facet_grid(rcp ~ .)
```

# Climate change simulation analysis

Variables that can be varied:

+ annual weather
  * balance (water deficit): irrigated and dryland yield 
  * days_ab_35: irrigated and dryland yield 
+ climate variable
  * balance average: share of irrigated acres 
  * days_ab_35 average: share of irrigated acres

Key limitations:

Water use is not modeled. As climate change progresses, more water is necessary.
However, its effect on saturated thickness is not modeled. Any simulations will not 
be realistic. Suppose we assume the historical trend of saturated thickness decline and use it
to project future saturated thickness level. This is not realistic since saturated thickness decline
should be faster because of greater water demand in future years than the past years.

Simulation variants:

+ **Irrigation share (fixed) and saturated thickness (fixed)**: 
This option answers the question of what would happen to average yield if farmers in the current 
production condition (irrigation share and saturated thickness) suddenly face weather like you will see
in the future.  
+ Irrigation share (varied based on climate) and saturated thickness (fixed): 
This option allows climate to influence irrigation share as climate change progresses. 
Under this option, irrigation share increases as the demand for irrigation increases. 
However, this ignores that fact that an increase in irrigation share has to reduce saturated 
thickness, which is assumed fixed.
+ Irrigation share (fixed) and saturated thickness (varied): inconsistent 
+ **Irrigation share (varied based on sat and climate) and saturated thickness (varied)**:

## Saturated thickness and ir_share fixed

```{r }
sim_ob <- 
  readRDS(here("Shared/Results/cl_simulationi_obs.rds")) %>% 
  .[, cl_period := case_when(
    real_year <= 2020 ~ "past",
    real_year > 2020 & real_year <= 2040 ~ "2021-2040",
    real_year > 2040 & real_year <= 2060 ~ "2041-2060",
    real_year > 2060 & real_year <= 2080 ~ "2061-2080",
    real_year > 2080 & real_year <= 2100 ~ "2081-2099",
  )] %>% 
  all_counties[., on = "sc_code"] %>% 
  nest_by(sc_code, crop_type) %>% 
  mutate(data_plot = list(
    data.table(data)[, .(y_ir_hat, y_dr_hat, avg_y_hat, cl_period, rcp)] %>% 
    melt(id.var = c("cl_period", "rcp")) %>% 
    .[, type := case_when(
      variable == "y_ir_hat" ~ "Irrigated Yield",
      variable == "y_dr_hat" ~ "Dryland Yield",
      variable == "avg_y_hat" ~ "Average Yield"
    )]
  )) %>% 
  mutate(g_cl_impact = list(
    ggplot(data_plot) +
      geom_boxplot(aes(y = value, x = cl_period, fill = factor(rcp))) +
      facet_grid(type ~ .) +
      ggtitle(
        paste0(
          "State = ",
          unique(data$state),
          ", County = ",
          unique(data$county_name),
          "\nSaturated Thickness = ", 
          round(unique(data$sat), digits = 2), 
          ", Share of Irrigated Acres = ", 
          round(unique(data$ir_share), digits = 3)
        )
      )
  ))
```

**Contributing Factors**

+ the impact of weather on irrigated yield (vary by saturate thickness)
+ the impact of weather on dryland yield
+ irrigated acreage share (fixed and not affected by saturated thickness)

By far the biggest contributor of the climate change impact of yield
is the second and third items, especially irrigated acreage share.
Irrigated yield can increase on average. Yield increases as balance increases
for a certain range (see the figures above). Depending on the change in the distribution of balance, 
climate change can increase yield on average.

### Corn

```{r }
filter(sim_ob, crop_type == "corn")$g_cl_impact
```

### Soy

```{r }
filter(sim_ob, crop_type == "soy")$g_cl_impact
```

## Average representative farm experiencing different kind of climate

Share of Irrigated Acres modeled for a few set of saturated thickness
levels, which are fixed for the entire period. Observed irrigated
share and saturated thickness level not respected.

Observations:
Share of irrigated acres goes up significantly because of warming climate without
any punishment on saturated thickness. Consequently, irrigated acres go up almost 
up to 1, leading to extreme resilience of average yield to climate change.

```{r }
sim_h <- 
  readRDS(here("Shared/Results/cl_simulationi_h.rds")) %>% 
  .[, cl_period := case_when(
    real_year <= 2020 ~ "past",
    real_year > 2020 & real_year <= 2040 ~ "2021-2040",
    real_year > 2040 & real_year <= 2060 ~ "2041-2060",
    real_year > 2060 & real_year <= 2080 ~ "2061-2080",
    real_year > 2080 & real_year <= 2100 ~ "2081-2099",
  )] %>% 
  all_counties[., on = "sc_code"] %>% 
  nest_by(sc_code, crop_type) %>% 
  mutate(data_plot = list(
    data.table(data)[, .(y_ir_hat, y_dr_hat, avg_y_hat, cl_period, rcp, sat, ir_share)] %>% 
    melt(id.var = c("cl_period", "rcp", "sat")) %>% 
    .[, type := case_when(
      variable == "y_ir_hat" ~ "Irrigated Yield",
      variable == "y_dr_hat" ~ "Dryland Yield",
      variable == "avg_y_hat" ~ "Average Yield",
      variable == "ir_share" ~ "Share of Irrigated Acres"
    )]
  )) %>% 
  mutate(g_cl_impact = list(
    ggplot(data_plot) +
      geom_boxplot(aes(y = value, x = cl_period, fill = factor(sat))) +
      facet_grid(type ~ rcp, scales = "free_y") +
      ggtitle(
        paste0(
          "State = ",
          unique(data$state),
          ", County = ",
          unique(data$county_name)
        )
      )
  ))

```
```{r eval = F, echo = F}
sim_h$g_cl_impact[[1]]
sim_h$data_plot[[1]]
data.table(sim_h$data[[1]])[, .(real_year, sat, ir_share)]
```

### Corn

```{r }
filter(sim_h, crop_type == "corn")$g_cl_impact
```

### Soy

```{r }
filter(sim_h, crop_type == "soy")$g_cl_impact
```