### 2023-07-21 Codes/Retired/3_2_make_figure_misc.Rmd

---
title: "Generate figures (climate, saturated thickness)"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Preparation

```{r }
#=== packages ===#
library(downloader)
library(terra)
library(lubridate)
library(tidyverse)
library(tidycensus)
library(data.table)
library(ncdf4)
library(sf)
library(raster)
library(parallel)
library(tmap)
library(here)
library(geosphere)
library(exactextractr)

#=== source functions ===#
source(here("GitControlled/Codes/functions.R"))

#=== Regression results ===#
# created in 1_regression_analysis.R
reg_results <- 
  here("Shared/Results", "reg_results.rds") %>% 
  readRDS() %>% 
  dplyr::select(crop_type, states, data, share_data, model_name) %>% 
  unique(by = "crop_type")
```

# Prepare Data

```{r echo = F}

#=== sf for yield regression ===#
sc_y_sf <-  
  reg_results %>% 
  dplyr::select(crop_type, data) %>% 
  unique(by = "crop_type") %>% 
  unnest() %>% 
  data.table() %>% 
  .[, .(sc_code, state, geometry, crop_type)] %>% 
  unique(by = c("sc_code", "crop_type"))

#=== sf for ir_share regression ===#
sc_is_sf <-  
  reg_results %>% 
  dplyr::select(crop_type, share_data) %>% 
  unnest() %>% 
  data.table() %>% 
  .[, .(sc_code, state, geometry, crop_type)] %>% 
  unique(by = c("sc_code", "crop_type"))

#=== hpa ===#
hpa_simplified <- 
  here("Shared/Data/ProcessedData/hp_simplified.shp") %>% 
  readRDS()

#=== all counties ===#
data(fips_codes)

all_counties <-
  tigris::counties() %>% 
  rename(
    state_code = STATEFP, 
    county_code = COUNTYFP,
    county_name = NAME
  ) %>% 
  left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
  dplyr::select(state_code, state, county_code, county_name) %>% 
  mutate(sc_code = paste0(state_code, county_code)) %>% 
  data.table()

all_states <- 
  tigris::states() %>% 
  rename(state_code = STATEFP) %>% 
  left_join(
    ., 
    unique(data.table(fips_codes), by = "state"), 
    by = "state_code"
  )

counties_states <- 
  reg_results %>%  
  mutate(all_counties = list(
    tigris::counties() %>% 
    rename(state_code = STATEFP, county_code = COUNTYFP) %>% 
    left_join(., fips_codes, by = c("state_code", "county_code")) %>% 
    filter(state %in% unique(share_data$state)) %>% 
    dplyr::select(state)  
  )) %>% 
  mutate(all_states = list(
    tigris::states() %>% 
    rename(state_code = STATEFP) %>% 
    left_join(
      ., 
      unique(data.table(fips_codes), by = "state"), 
      by = "state_code"
    ) %>% 
    filter(state %in% unique(share_data$state)) %>% 
    dplyr::select(state) 
  ))
```

# Geographic Focus
## Yield regression
Map of counties used for yield regression

+ Irrigated yield observations outside of HPA are dropped because saturated thickness is not observed
+ Dryland yield observations are kept even if outside of HPA
+ Both dryland and irrigated yield observations are use if inside of HPA

```{r }
states_y <- 
  sc_y_sf %>% 
  nest_by(crop_type) %>% 
  mutate(state_ls = list(
    unique(data$state)
  )) %>% 
  mutate(state_sf = list(
    filter(all_states, state %in% state_ls)
  )) %>% 
  dplyr::select(crop_type, state_sf) %>% 
  unnest()

ggplot() + 
  geom_sf(data = st_as_sf(sc_y_sf), fill = "green", alpha = 0.4) +
  geom_sf(data = st_as_sf(states_y), color = "red", fill = NA) +
  geom_sf(data = hpa_simplified, fill = "blue", alpha = 0.2) +
  facet_grid(. ~ crop_type)
```

## Irrigation Share (and Simulation)
Map of counties used for irrigation share regression and simulations

+ Only the counties with 95% of its area over HPA are kept

```{r }
states_is <- 
  sc_is_sf %>% 
  nest_by(crop_type) %>% 
  mutate(state_ls = list(
    unique(data$state)
  )) %>% 
  mutate(state_sf = list(
    filter(all_states, state %in% state_ls)
  )) %>% 
  dplyr::select(crop_type, state_sf) %>% 
  unnest()

ggplot() + 
  geom_sf(data = st_as_sf(sc_is_sf), fill = "green", alpha = 0.4) +
  geom_sf(data = st_as_sf(states_is), color = "red", fill = NA) +
  geom_sf(data = hpa_simplified, fill = "blue", alpha = 0.2) +
  facet_grid(. ~ crop_type)
```

# Saturated Thickness

```{r }
sat_sf <- 
  reg_results %>% 
  dplyr::select(crop_type, share_data) %>% 
  unique(by = "crop_type") %>% 
  unnest() %>% 
  nest_by(sc_code, crop_type) %>% 
  #=== sat trend ===#
  mutate(beta_sat = 
    lm(sat ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== average sat level ===#
  mutate(sat_mean = 
    mean(data$sat, na.rm = TRUE)
  ) %>% 
  mutate(geometry = 
    unique(data$geometry)
  ) %>% 
  st_as_sf() 
```

## Saturated thickness trend by sc_code
`beta_sat` measures the annual rate of change in saturated thickness
observed 1990 - 2016

```{r hist-sat-trend, fig.cap = "Saturated thickness trend (1990-2016)"}
ggplot() +
  geom_histogram(
    data = sat_sf, 
    aes(x = beta_sat), 
    color = "blue",
    fill = NA
  ) +
  facet_grid(crop_type ~ .)
```

## Histogram of saturated thickness level by sc_code

```{r hist-sat-level, fig.cap = "Mean saturated thickness (1990-2016)"}
ggplot() +
  geom_histogram(
    data = sat_sf, 
    aes(x = sat_mean), 
    color = "blue",
    fill = NA,
    bins = 50
  ) +
  facet_grid(crop_type ~ .)
```

## Histogram of saturated thickness level (below 100)

```{r hist-sat-level-b100, fig.cap = "Mean saturated thickness (less than 100) (1990-2018)"}
ggplot() +
  geom_histogram(
    data = filter(sat_sf, sat_mean <= 100), 
    aes(x = sat_mean), 
    color = "blue",
    fill = NA,
    bins = 50
  ) +
  facet_grid(crop_type ~ .)
```

## Map of saturated thickness level 
Counties with fill colors are used for simulation and ir-share regression. 
Corn: NE, KS, CO, SD 
Soy: NE, KS

```{r }
sc_is_sf <-
  left_join(sc_is_sf, dplyr::select(sat_sf, sc_code, beta_sat, sat_mean), by = "sc_code")

```
```{r hist-sat-map, fig.cap = "Map of mean saturated thickness (1990-2018)"}
ggplot() +
geom_sf(data = st_as_sf(states_is), fill = NA, col = "red") + 
geom_sf(data = hpa_simplified, fill = "blue", alpha = 0.4) + 
geom_sf(data = st_as_sf(sc_is_sf), aes(fill = sat_mean)) + 
facet_grid(. ~ crop_type) +
scale_fill_viridis_c()
```

# Basic data visualization

```{r }
data_y <- reg_results %>% 
  dplyr::select(crop_type, data) %>% 
  unique(by = "crop_type") %>% 
  unnest()
```

## Yield histogram by saturated thickness category

```{r yield-hist}
ggplot(data_y) +
  geom_boxplot(aes(y = yield, x = sat_cat)) +
  facet_grid(crop_type ~ ., scales = "free_y")
```

## Yield against water deficit (balance)

```{r yield-balance}
ggplot(data_y) +
  geom_point(aes(y = yield, x = balance, color = sat_cat), size = 0.4) +
  geom_smooth(aes(y = yield, x = balance, color = sat_cat)) +
  facet_grid(crop_type ~ ., scales = "free_y")
```

# Climate data 
## Prepare data

```{r }
model_ls <- 
  c(
    "GFDL-ESM2M", 
    "HadGEM2-ES365", 
    "IPSL-CM5A-LR", 
    "MIROC5"
  )

rcp_ls <- c(45, 85)

model_rcp <- 
  expand.grid(
    cl_model = model_ls,
    rcp = rcp_ls
  ) %>% 
  data.table()

#=== weather data in the data for regression ===#
# cl_data_current <- 
#   reg_data[, .(sc_code, year, balance, et0, prcp, days_ab_35)] %>% 
#   expand_grid_df(., model_rcp)

#=== future climate data ===#
cl_data_future <- 
  model_rcp %>% 
  rowwise() %>% 
  #=== define maca data file names ===#
  mutate(file_name = list(
    paste0(
      "Shared/Data/ProcessedData/MACA/sum_", 
      cl_model,
      "_",
      rcp,
      "_county.rds"
    ) %>% 
    here()
  )) %>% 
  mutate(maca_data = list(
    #=== read maca data ===#
    readRDS(file_name) %>% 
    #=== calculate 10-year balance avg ===#
    .[, 
      `:=`(
        balance_avg = mean(balance), 
        days_ab_35_avg = mean(days_ab_35)
      ), 
      by = .(floor((year - 2020) / 10), sc_code)
    ] 
  )) %>% 
  dplyr::select(cl_model, rcp, maca_data) %>% 
  unnest() %>% 
  data.table() %>% 
  .[, .(sc_code, year, balance, et0, prcp, days_ab_35, cl_model, rcp)]  %>% 
  .[, cl_period := case_when(
    year <= 2020 ~ "past",
    year > 2020 & year <= 2040 ~ "2021-2040",
    year > 2040 & year <= 2060 ~ "2041-2060",
    year > 2060 & year <= 2080 ~ "2061-2080",
    year > 2080 & year <= 2100 ~ "2081-2099",
  )] %>% 
  all_counties[., on = "sc_code"]  
```

## Weather variables by period (model mixed)

```{r cl-trend-future, fig.cap = "Water deficit, ET0, and precipitation by period (model mixed)"}
cl_data_future[, .(balance, et0, prcp, days_ab_35, rcp, sc_code, year, cl_period)] %>% 
melt(id.var = c("sc_code", "year", "rcp", "cl_period")) %>% 
ggplot(data = ) +
  geom_density(aes(x = value, fill = cl_period), alpha = 0.4) +
  facet_grid(rcp ~ variable, scales = "free") +
  theme(
    legend.position = "bottom"
  )
```

## Rate of increase in water deficit by county and model

```{r }
cl_trend_county <- 
  cl_data_future %>% 
  nest_by(sc_code, cl_model, rcp) %>% 
  #=== rate of balance increase ===#
  mutate(beta_balance = 
    lm(balance ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== rate of et0 increase ===#
  mutate(beta_et0 = 
    lm(et0 ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== rate of precip increase ===#
  mutate(beta_prcp = 
    lm(prcp ~ year, data = data)$coef[["year"]]
  ) %>% 
  #=== rate of precip increase ===#
  mutate(beta_d35 = 
    lm(days_ab_35 ~ year, data = data)$coef[["year"]]
  ) %>% 
  mutate(geometry = 
    unique(data$geometry)
  ) %>% 
  st_as_sf()
```

### Balance

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_balance)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### Precipitation

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_prcp)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### ET0

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_et0)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```

### days_above_35

```{r }
ggplot(cl_trend_county) +
  geom_sf(aes(fill = beta_d35)) + 
  facet_grid(cl_model ~ rcp) +
  scale_fill_viridis_c()
```