### 2023-07-22 Codes/0_1_hpa_counties.Rmd
Further revisions:
1. confidence intervals for all
2. testing the impact of aquifer thickness
3. specific aquifer thickness values for average yield analysis

---
title: "Prepare base-county data"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Identify counties the overlaps with the HPA boundary
## Read the HPA border shapefile

```{r Read-HPA, cache = TRUE}
hpa <- st_read(here("Shared/Data/RawData/hp_bound2010.shp"))
```

Map-of-HPA

```{r }
tm_shape(st_simplify(hpa)) +
  tm_polygons()
```

## Get all the US counties as an sf

```{r Get-US-counties}
(
US_county <- tigris::counties(cb = TRUE) %>%
  st_as_sf() %>%
  st_transform(st_crs(hpa)) %>%
  rename("state_code" = "STATEFP", "county_code" = "COUNTYFP") %>%
  dplyr::select(state_code, county_code)
)
```

## Identify counties that overlaps with `hpa`

```{r Identify-overlap, cache = TRUE}
hpa_counties <- US_county[hpa, ] %>%
  mutate(whole_area = st_area(.) %>% as.numeric())
```

Map-of-HPA and the intersecting counties

```{r }
tm_shape(st_simplify(hpa)) +
  tm_polygons() +
tm_shape(hpa_counties) +
  tm_polygons(col = "blue", alpha = 0.4)
```

## Find the area of the portion that is intersecting with hpa
Intersect HPA and counties in HPA

```{r intersect-HPA-county}
intersected_portion <- st_intersection(hpa, hpa_counties) %>%
  #=== get the intersecting area ===#
  mutate(area = st_area(.) %>% as.numeric()) %>%
  data.table() %>%
  .[, .(interesected_area = sum(area)), by = .(state_code, county_code)]
```

Merge the intersection data

```{r }
hpa_counties <- 
  left_join(
    hpa_counties,
    intersected_portion,
    by = c("state_code", "county_code")
  ) %>%
  #=== find the proportion of the county area that overlaps with HPA ===#
  mutate(ir_area_ratio = interesected_area / whole_area) %>%
  data.table() %>%
  .[, geometry := NULL]
```

Histogram of coverage ratio

```{r }
hpa_counties$ir_area_ratio %>% hist
```

Get the list of the states

```{r }
hpa_states <- hpa_counties$state_code %>% unique()
```

# Additional counties 
Additional state codes:
these supplement the dryland production data for understanding the impact of drought severity
on dryland crop production

```{r }
data(fips_codes)

additional_states <- fips_codes %>%
  data.table() %>%
  unique(by = "state_name") %>%
  .[state_name %in% c("Illinois", "Iowa", "Missouri", "Indiana"), ] %>%
  .[, state_code]

all_states <- c(hpa_states, additional_states)
```

# All counties
**Get the state abbreviation names**

```{r }
fips_codes <- fips_codes %>%
  data.table() %>%
  .[, .(state, state_code, state_name, county_code)]
```

Get all the counties in the target states

```{r }
(
all_counties <- filter(US_county, state_code %in% all_states) %>%
  data.table() %>%
  #=== merge with the data on counties in HPA ===#
  hpa_counties[., on = c("state_code", "county_code")] %>%
  #=== merge with FIPS codes data ===#
  fips_codes[., on = c("state_code", "county_code")] %>%
  #=== indicator of whether intersecting with HPA or not ===#
  .[, in_hpa := 0] %>%
  .[!is.na(ir_area_ratio), in_hpa := 1] %>%
  #=== generate state-code variable ===#
  .[, sc_code := paste0(state_code, county_code)] %>%
  #=== keep only the relevant variables ===#
  .[, .(state, state_name, sc_code, ir_area_ratio, geometry, in_hpa)]
)
```

Map of all the counties: 
Note: not all the counties are used in our eventual analysis

```{r map-all-counties, cache = TRUE}
tm_shape(st_as_sf(all_counties)) +
tm_polygons() +
tm_shape(
  tigris::states() %>% 
    filter(STATEFP %in% all_states)
) +
tm_borders(col = "red")
```

# save the county data

```{r }
saveRDS(all_counties, here("Shared/Data/ProcessedData/base_counties.rds"))

# all_counties[, state_name] %>% unique()
```