### 2023-07-22 Codes/0_3_get_sat.Rmd

---
title: "Assign saturated thickness value "
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi=400,
  fig.width=7.5,
  fig.height=5,
  out.width="750px",
  out.height="500px"
)
```

# Preparation
Import saturated thickness data

```{r }
sat_temp <- raster("/Volumes/Elements/Database/SaturatedThickness/SatThick_19851.tif")
```

HPA counties

```{r }
hpa_counties <- 
  here("Shared/Data/ProcessedData/base_counties.rds") %>% 
  readRDS() %>% 
  .[in_hpa == 1, ] %>% 
  st_as_sf() %>% 
  st_transform(projection(sat_temp)) 
```

Define function to extract saturated thickness for each county

```{r }
get_sat <- function(y) {
  # y <- 2000
  sat_temp <- 
    paste0("/Volumes/Elements/Database/SaturatedThickness/SatThick_", y, "1.tif") %>% 
    raster()

  sat_mean <- 
    exact_extract(sat_temp, hpa_counties, "mean")

  data_return <- 
    data.table(
      sat = sat_mean,
      sc_code = hpa_counties$sc_code,
      year = y
    )

  return(data_return)
}
```

Extract saturated thickness values (parallelized)

```{r extract-sat, eval = F}
sat_all <- mclapply(1985:2016, get_sat, mc.cores = 12) %>%
  rbindlist() %>%
  .[sat < 0, sat := 0] %>% 
  .[, sat := sat * 3.28084] # meter to feet
```

# Save the data

```{r save, eval = F}
saveRDS(sat_all, here("Shared/Data/ProcessedData/sat_all.rds"))
```

# Quick visualization

```{r }
sat_all <- 
  here("Shared/Data/ProcessedData/sat_all.rds") %>% 
  readRDS()
```

Saturate thickness history by county

```{r }
ggplot(sat_all) +
  geom_line(aes(y = sat, x = year, group = sc_code))
```

Histogram

```{r }
ggplot(sat_all) +
  geom_histogram(aes(x = sat))
```