### 2023-07-22 Codes/1_2_test_dif_significance_ir_yield.rmd

---
title: Test of the significance of yield difference (intensive)
---

# Preparation

```{r}
results <- readRDS("Shared/Results/reg_results_data.rds")
```

# Define the function to test

```{r}
test_dif_in_yield <- function(crop, balance, base_q, comp_q, results) {
  #++++++++++++++++++++++++++++++++++++
  #+ For testing
  #++++++++++++++++++++++++++++++++++++
  # crop <- "Corn"
  # balance <- 1000
  # base_q <- 1
  # comp_q <- 3

  #++++++++++++++++++++++++++++++++++++
  #+ Main
  #++++++++++++++++++++++++++++++++++++
  q_data <-
    data.table(
      quater = c(1, 2, 3),
      coef_name = list(
        c("s_balance_sat_cat_39_4_82__1", "s_balance_sat_cat_39_4_82__2", "sat_cat::[39.4,82]"),
        c("s_balance_sat_cat_82_194__1", "s_balance_sat_cat_82_194__2", "sat_cat::(82,194]"),
        c("s_balance_sat_cat_194_708__1", "s_balance_sat_cat_194_708__2", "sat_cat::(194,708]")
      ),
      quarter_names = c("[39.4,82]", "(82,194]", "(194,708]")
    )

  crop_w <- crop
  results_w <- results[crop == crop_w, ]
  fe_res <- results_w[, semi_res][[1]]$fe_res
  vcov <- vcov(fe_res)
  gam_res <- results_w[, semi_res][[1]]$gam_res
  data_w <- results_w[, data][[1]]
  balance_w <- balance

  #--- create data for prediction ---# 
  data_for_pred <-
    CJ(
      balance = balance_w,
      sat_cat = unique(data_w$sat_cat)
    ) %>%
    # === can be any sc_code (need to shift yield for "average" county) ===#
    # .[, sc_code := data_w$sc_code[2]] %>%
    .[, sc_code := data_w[sat_cat != "dryland"]$sc_code[1]] %>%
    .[, sc_dry := paste0(data_w$sc_dry[1])] %>%
    # === can be any year (need to shift yield for "average" year) ===#
    .[, year := 2009] %>%
    # === fake yield data (necessary to take advantage of predict.gam) ===#
    .[, yield := 0] %>%
    .[, type := "pred"] %>%
    .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0) %>% factor()]

  #--- create bases  ---#
  yhat_data <-
    gen_smooth_data(
      data = data_for_pred,
      gam_res = gam_res
    ) %>%
    .$data

  #++++++++++++++++++++++++++++++++++++
  #+ point estimate
  #++++++++++++++++++++++++++++++++++++
  coef_data <- tidy(fe_res) %>% data.table()

  #--- coefficients ---#
  base_coef_names <- q_data[quater == base_q, coef_name][[1]]
  comp_coef_names <- q_data[quater == comp_q, coef_name][[1]]

  base_coefs <- coef_data[term %in% base_coef_names, estimate]
  comp_coefs <- coef_data[term %in% comp_coef_names, estimate]

  coefs <- c(comp_coefs, base_coefs)

  #--- X ---#
  base_x_names <- base_coef_names[-3]
  comp_x_names <- comp_coef_names[-3]

  base_x <-
    c(
      yhat_data[balance == balance_w & sat_cat == q_data[quater == base_q, quarter_names], ..base_x_names] %>% unlist(),
      1
    )

  comp_x <-
    c(
      yhat_data[balance == balance_w & sat_cat == q_data[quater == comp_q, quarter_names], ..comp_x_names] %>% unlist(),
      1
    )

  x_vec <- c(comp_x, -base_x)

  #--- difference in yield ---#
  dif_hat <- sum(coefs * x_vec)

  #++++++++++++++++++++++++++++++++++++
  #+ SE
  #++++++++++++++++++++++++++++++++++++
  coef_names <- c(base_coef_names, comp_coef_names)
  which_col <- purrr::map_dbl(coef_names, function(x) which(x == colnames(vcov)))
  rel_vcov <- vcov[which_col, which_col]

  se <- sqrt(t(x_vec) %*% rel_vcov %*% x_vec) %>% .[1, 1]

  test_result <-
    data.table(
      crop = crop_w,
      balance = balance_w,
      base_q = base_q,
      comp_q = comp_q,
      dif_hat = dif_hat,
      se = se,
      t = dif_hat / se
    )

  return(test_result)
}
```

# Run tests for a series of values of water deficit

```{r}
data_for_test <-
  expand.grid(
    crop = c("Corn", "Soybean"),
    balance = seq(-200, 1200, by = 20),
    comp_q = c(2, 3)
  ) %>% data.table()

test_results <-
  lapply(
    1:nrow(data_for_test),
    function(x) {
      test_dif_in_yield(
        crop = data_for_test[x, crop],
        balance = data_for_test[x, balance],
        base_q = 1,
        comp_q = data_for_test[x, comp_q],
        results = results
      )
    }
  ) %>%
  rbindlist()

saveRDS(test_results, "Shared/Results/test_results_intensive.rds")
```