### 2023-07-29 Codes/1_Analysis.rmd

---
title: "Regression Analysis and Prediction"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Prepare data

```{r}
#--- read data ---#
final_data <-
  here("Shared/Data/ProcessedData/final_data.rds") %>%
  # here::here("../Shared/Data/ProcessedData/final_data.rds") %>%
  readRDS()

#--- Data used for regression analysis ---#
reg_data <-
  prepare_reg_data(
    data = final_data,
    sat_thld_m = 9, # saturated thickness has to be at least 12 meters
    ir_share_thld = 0.75, # at least 75% of total county area has to be overlapped with HPA
    balance_thld = c(-1200, 200),
    sat_cat_num = 3,
    state_ls =
      list(
        corn = c("CO", "KS", "NE", "NM", "SD", "TX", "WY"),
        soy = c("KS", "NE", "TX")
      )
  ) %>%
  data.table() %>%
  setnames("crop_type", "crop") %>%
  #--- further necessary transformation ---#
  dplyr::mutate(data = purrr::map(
    reg_data_y,
    \(reg_data_y) {
      copy(reg_data_y) %>%
        #--- bu/acre to tonne/ha ---#
        .[, yield := yield * 12.553 / 200] %>%
        .[, dry_or_not := fifelse(sat_cat == "dryland", 1, 0)] %>%
        .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
        .[, state_year := paste0(state_name, "_", year)] %>%
        .[, balance := -balance]
    }
  ))

# saveRDS(reg_data, file = here::here("Shared/Data/reg_data.rds"))

# ggplot(reg_data$data[[1]]) +
#   geom_point(aes(y = acres_ratio, x = sat)) +
#   geom_smooth(aes(y = acres_ratio, x = sat))
```

# Regression and prediciton

## Yield

```{r}
set.seed(283397)

#--- global parameters ---#
sc_base <- "31011"
sc_dry_base <- "31011_0"

main_analysis <-
  reg_data %>%
  rowwise() %>%
  #--- the same sequence used for all the bootstrap iterations ---#
  dplyr::mutate(balance_seq = list(
    seq(
      min(data$balance),
      max(data$balance),
      length = 100
    )
  )) %>%
  dplyr::mutate(sat_seq_eval = list(
    c(10, 50, 90, 130)
  )) %>%
  dplyr::mutate(share_reg_data = list(
    data %>%
      .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
      .[, balance_avg := mean(balance), by = sc_code]
  )) %>%
  #--- the same sequence used for all the bootstrap iterations ---#
  dplyr::mutate(sat_seq = list(
    share_reg_data[, seq(min(sat),
      max(sat),
      length = 100
    )] %>%
      c(., sat_seq_eval)
  )) %>%
  #--- values of the controls to be used in prediction ---#
  dplyr::mutate(
    sandtotal_med = median(share_reg_data$sandtotal_r),
    silttotal_med = median(share_reg_data$silttotal_r),
    awc_med = median(share_reg_data$awc_r)
  ) %>%
  #++++++++++++++++++++++++++++++++++++
  #+ Main analysis
  #++++++++++++++++++++++++++++++++++++
  dplyr::mutate(yield_pred_data = list(
    yield_analysis(data, balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
  )) %>%
  dplyr::mutate(share_pred_data = list(
    share_analysis(share_reg_data, sat_seq, sandtotal_med, silttotal_med, awc_med)
  )) %>%
  dplyr::mutate(avg_yield_pred_data = list(
    get_average_yield(yield_pred_data, share_pred_data)
  ))
```

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Bootstrap
#++++++++++++++++++++++++++++++++++++
boot_results_added <-
  main_analysis %>%
  #--- bootsrap data ---#
  dplyr::mutate(boot_data = list(
    replicate(1000, boot(data), simplify = FALSE)
  )) %>%
  #--- get share data ---#
  dplyr::mutate(share_boot_data = list(
    lapply(
      1:length(boot_data),
      \(x) {
        boot_data[[x]] %>%
          .[ir == "ir" & ir_area_ratio >= 0.75, ] %>%
          .[, balance_avg := mean(balance), by = sc_code]
      }
    )
  )) %>%
  #--- yield analysis ---#
  dplyr::mutate(yield_pred_data_boot = list(
    parallel::mclapply(
      1:length(boot_data),
      \(x) {
        print(x)
        yield_analysis(boot_data[[x]], balance_seq, sat_seq_eval, sat_breaks, sat_text_data)
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  #--- share analysis ---#
  dplyr::mutate(share_pred_data_boot = list(
    parallel::mclapply(
      1:length(share_boot_data),
      \(x) {
        print(x)
        share_analysis(share_boot_data[[x]], sat_seq, sandtotal_med, silttotal_med, awc_med)
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  #--- average yield analysis ---#
  dplyr::mutate(avg_yield_pred_data_boot = list(
    parallel::mclapply(
      1:length(boot_data),
      \(x) {
        get_average_yield(yield_pred_data_boot[[x]], share_pred_data_boot[[x]])
      },
      mc.cores = parallel::detectCores() - 2
    )
  )) %>%
  dplyr::mutate(
    yield_pred_data_boot = list(rbindlist(yield_pred_data_boot)),
    share_pred_data_boot = list(rbindlist(share_pred_data_boot)),
    avg_yield_pred_data_boot = list(rbindlist(avg_yield_pred_data_boot))
  ) %>%
  dplyr::select(
    crop, data, sat_seq, sat_text_data,
    yield_pred_data, yield_pred_data_boot,
    share_pred_data, share_pred_data_boot,
    avg_yield_pred_data, avg_yield_pred_data_boot,
  )

saveRDS(boot_results_added, "Shared/Results/boot_results_added.rds")
```

```{r}
#++++++++++++++++++++++++++++++++++++
#+ Summarize
#++++++++++++++++++++++++++++++++++++
boot_results_added <- readRDS("Shared/Results/boot_results_added.rds")

all_results <-
  boot_results_added %>%
  # rowwise() %>%
  dplyr::mutate(yield_response_sum = list(
    yield_pred_data_boot[, .(
      # y_hat_boot = median(y_hat, na.rm = TRUE),
      y_hat_se = sd(y_hat, na.rm = TRUE),
      # dif_y_hat_boot = mean(dif_y_hat, na.rm = TRUE),
      dif_y_hat_se = sd(dif_y_hat, na.rm = TRUE)
    ),
    by = .(balance, sat, sat_cat)
    ] %>%
      .[yield_pred_data, on = c("balance", "sat")] %>%
      .[, `:=`(
        lower = y_hat - 1.96 * y_hat_se,
        upper = y_hat + 1.96 * y_hat_se,
        dif_lower = dif_y_hat - 1.96 * dif_y_hat_se,
        dif_upper = dif_y_hat + 1.96 * dif_y_hat_se
      )]
  )) %>%
  dplyr::mutate(share_sum = list(
    share_pred_data_boot[, .(
      # ir_share_hat = mean(ir_share_hat, na.rm = TRUE),
      ir_share_hat_se = sd(ir_share_hat, na.rm = TRUE)
    ),
    by = .(sat)
    ] %>%
      .[share_pred_data, on = "sat"] %>%
      .[, `:=`(
        lower = ir_share_hat - 1.96 * ir_share_hat_se,
        upper = ir_share_hat + 1.96 * ir_share_hat_se
      )]
  )) %>%
  dplyr::mutate(avg_yield_sum = list(
    avg_yield_pred_data_boot[, .(
      avg_yield_se = sd(avg_yield, na.rm = TRUE),
      dif_avg_yield_se = sd(dif_avg_yield, na.rm = TRUE)
    ),
    by = .(balance, sat)
    ] %>%
      .[avg_yield_pred_data, on = c("balance", "sat")] %>%
      .[, `:=`(
        upper = avg_yield + 1.96 * avg_yield_se,
        lower = avg_yield - 1.96 * avg_yield_se,
        dif_upper = dif_avg_yield + 1.96 * dif_avg_yield_se,
        dif_lower = dif_avg_yield - 1.96 * dif_avg_yield_se
      )]
  )) %>%
  data.table()

saveRDS(all_results, "Shared/Results/all_results.rds")
```