### 2023-07-29 Codes/1_2_test_dif_significance_ir_yield.rmd

---
title: Test of the significance of yield difference (intensive)
---

# Preparation

```{r}
results <- readRDS("Shared/Results/reg_results_data.rds")
```

```{r}
main_analysis <- readRDS("Shared/Results/main_analysis.rds")

semi_res <- main_analysis$yield_semi_res[[1]]
sat_text_data <- main_analysis$sat_text_data[[1]]
```

# Define the function to test

```{r}
main_analysis <-
  main_analysis %>%
  dplyr::mutate(yield_dif_test = list(
    lapply(
      2:3,
      \(x) {
        test_dif_in_yield(
          balance_ls = seq(min(data$balance), max(data$balance), length = 50),
          base_q = 1,
          comp_q = x,
          yield_semi_res,
          sat_text_data,
          sc_base
        ) %>%
        .[, comp_q := x]
      }
    ) %>%
    rbindlist()
  ))

test_dif_in_yield <- function(balance_ls, base_q, comp_q, yield_semi_res, sat_text_data, sc_base) {
  #++++++++++++++++++++++++++++++++++++
  #+ For testing
  #++++++++++++++++++++++++++++++++++++
  # crop_w <- "corn"
  # balance <- 1000
  # base_q <- 1
  # comp_q <- 3

  # balance_ls <- c(200, 300, 400)

  #++++++++++++++++++++++++++++++++++++
  #+ Main
  #++++++++++++++++++++++++++++++++++++
  sc_dry_0 <- paste0(sc_base, "_1")
  sc_dry_ir <- paste0(sc_base, "_0")

  q_data <-
    sat_text_data %>%
    .[, coef_name := paste0("s_balance_sat_cat", sat_cat)] %>%
    .[, coef_name := gsub("\\[", "_", coef_name)] %>%
    .[, coef_name := gsub("\\]", "", coef_name)] %>%
    .[, coef_name := gsub("\\(", "_", coef_name)] %>%
    .[, coef_name := gsub("\\,", "_", coef_name)] %>%
    .[, coef_name := gsub("\\.", "_", coef_name)] %>%
    rowwise() %>%
    dplyr::mutate(coef_name = list(
      c(
        paste0(coef_name, "__", c(1, 2)),
        paste0("sat_cat::", sat_cat)
      )
    )) %>%
    data.table()

  fe_res <- yield_semi_res$fe_res
  vcov <- vcov(fe_res)
  gam_res <- yield_semi_res$gam_res

  #--- create data for prediction ---#
  data_for_pred <-
    CJ(
      balance = balance_ls,
      sat_cat = unique(q_data$sat_cat)
    ) %>%
    # === can be any sc_code (need to shift yield for "average" county) ===#
    # .[, sc_code := data_w$sc_code[2]] %>%
    .[, sc_dry := sc_dry_ir] %>%
    # === can be any year (need to shift yield for "average" year) ===#
    .[, year := 2009] %>%
    # === fake yield data (necessary to take advantage of predict.gam) ===#
    .[, yield := 0]

  #--- create bases  ---#
  yhat_data <-
    gen_smooth_data(
      data = data_for_pred,
      gam_res = gam_res
    ) %>%
    .$data

  #++++++++++++++++++++++++++++++++++++
  #+ Coefficients
  #++++++++++++++++++++++++++++++++++++
  coef_data <- tidy(fe_res) %>% data.table()

  #--- coefficients ---#
  base_coef_names <- q_data[sat_rank == base_q, coef_name][[1]]
  comp_coef_names <- q_data[sat_rank == comp_q, coef_name][[1]]

  base_coefs <- coef_data[term %in% base_coef_names, estimate]
  comp_coefs <- coef_data[term %in% comp_coef_names, estimate]

  coefs <- c(comp_coefs, base_coefs)

  #--- X ---#
  base_x_names <- base_coef_names[-3]
  comp_x_names <- comp_coef_names[-3]

  #++++++++++++++++++++++++++++++++++++
  #+ loop over balance
  #++++++++++++++++++++++++++++++++++++
  lapply(
    1:length(balance_ls),
    \(x) {
      balance_w <- balance_ls[x]
      #---------------------
      #- point estimate
      #---------------------
      base_x <-
        c(
          yhat_data[balance == balance_w & sat_cat == q_data[sat_rank == base_q, sat_cat], ..base_x_names] %>% unlist(),
          1
        )

      comp_x <-
        c(
          yhat_data[balance == balance_w & sat_cat == q_data[sat_rank == comp_q, sat_cat], ..comp_x_names] %>% unlist(),
          1
        )

      x_vec <- c(comp_x, -base_x)

      #--- difference in yield ---#
      dif_hat <- sum(coefs * x_vec)

      #++++++++++++++++++++++++++++++++++++
      #+ SE
      #++++++++++++++++++++++++++++++++++++
      coef_names <- c(base_coef_names, comp_coef_names)
      which_col <- purrr::map_dbl(coef_names, function(x) which(x == colnames(vcov)))
      rel_vcov <- vcov[which_col, which_col]

      se <- sqrt(t(x_vec) %*% rel_vcov %*% x_vec) %>% .[1, 1]

      test_result <-
        data.table(
          balance = balance_w,
          base_q = base_q,
          comp_q = comp_q,
          dif_hat = dif_hat,
          se = se,
          t = dif_hat / se
        )

      return(test_result)
    }
  ) %>%
    rbindlist()
}

```

# Run tests for a series of values of water deficit

```{r}
data_for_test <-
  expand.grid(
    crop = c("Corn", "Soybean"),
    balance = seq(-200, 1200, by = 20),
    comp_q = c(2, 3)
  ) %>% data.table()

test_results <-
  lapply(
    1:nrow(data_for_test),
    function(x) {
      test_dif_in_yield(
        crop = data_for_test[x, crop],
        balance = data_for_test[x, balance],
        base_q = 1,
        comp_q = data_for_test[x, comp_q],
        results = results
      )
    }
  ) %>%
  rbindlist()

saveRDS(test_results, "Shared/Results/test_results_intensive.rds")
```