### 2023-08-30 Codes/0_2_quickstat_wrangle.rmd
minor updates

```{r}
yields <-
  fread("Shared/Data/RawData/USDA-NASS/yields.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  setnames(
    c("county ansi", "state ansi", "data item", "value", "commodity"),
    c("county_fips", "state_fips", "data_item", "yield", "crop")
  ) %>%
  .[!str_detect(data_item, "SILAGE"), ] %>%
  .[!is.na(county_fips), ] %>%
  .[, crop := tolower(crop)] %>%
  .[, crop := ifelse(crop == "soybeans", "soy", "corn")] %>%
  .[, yield_type := case_when(
    str_detect(data_item, "NET") ~ "yield_npa",
    !str_detect(data_item, "NET") ~ "yield_pa"
  )] %>%
  .[, ir := ifelse(str_detect(data_item, "NON"), "nir", "ir")] %>%
  .[, county_fips := str_pad(county_fips, 3, "left", "0")] %>%
  .[, state_fips := str_pad(state_fips, 2, "left", "0")] %>%
  .[, sc_code := paste0(state_fips, county_fips)] %>%
  .[, .(sc_code, year, crop, ir, yield, yield_type)] %>%
  dcast(sc_code + crop + year + ir ~ yield_type, value.var = "yield")

harvested_acres <-
  fread("Shared/Data/RawData/USDA-NASS/harvested_acres.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  setnames(
    c("county ansi", "state ansi", "data item", "value", "commodity"),
    c("county_fips", "state_fips", "data_item", "acres", "crop")
  ) %>%
  .[!str_detect(data_item, "SILAGE"), ] %>%
  .[!is.na(county_fips), ] %>%
  .[, crop := tolower(crop)] %>%
  .[, crop := ifelse(crop == "soybeans", "soy", "corn")] %>%
  .[, county_fips := str_pad(county_fips, 3, "left", "0")] %>%
  .[, state_fips := str_pad(state_fips, 2, "left", "0")] %>%
  .[, sc_code := paste0(state_fips, county_fips)] %>%
  .[, ir := ifelse(str_detect(data_item, "NON"), "nir", "ir")] %>%
  .[, harvested_acres := as.numeric(gsub(",", "", acres))] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(crop, sc_code, year)] %>%
  .[comp == 2, ] %>%
  .[, acres := as.numeric(gsub(",", "", acres))] %>%
  .[, acres_ratio_h := acres / sum(acres), by = .(crop, sc_code, year)] %>%
  .[, .(sc_code, year, crop, ir, harvested_acres, acres_ratio_h)]

planted_acres <-
  fread("Shared/Data/RawData/USDA-NASS/planted_acres.csv") %>%
  setnames(names(.), tolower(names(.))) %>%
  setnames(
    c("county ansi", "state ansi", "data item", "value", "commodity"),
    c("county_fips", "state_fips", "data_item", "acres", "crop")
  ) %>%
  .[!str_detect(data_item, "SILAGE"), ] %>%
  .[!is.na(county_fips), ] %>%
  .[, crop := tolower(crop)] %>%
  .[, crop := ifelse(crop == "soybeans", "soy", "corn")] %>%
  .[, county_fips := str_pad(county_fips, 3, "left", "0")] %>%
  .[, state_fips := str_pad(state_fips, 2, "left", "0")] %>%
  .[, sc_code := paste0(state_fips, county_fips)] %>%
  .[, ir := ifelse(str_detect(data_item, "NON"), "nir", "ir")] %>%
  .[, planted_acres := as.numeric(gsub(",", "", acres))] %>%
  .[, comp := sum(ir == "nir" | ir == "ir"), by = .(crop, sc_code, year)] %>%
  .[comp == 2, ] %>%
  .[, acres := as.numeric(gsub(",", "", acres))] %>%
  .[, acres_ratio_p := acres / sum(acres), by = .(crop, sc_code, year)] %>%
  .[, .(sc_code, year, crop, ir, planted_acres, acres_ratio_p)]
```

```{r}
nass_data_new <-
  harvested_acres[yields, on = c("sc_code", "ir", "year", "crop")] %>%
  planted_acres[., on = c("sc_code", "ir", "year", "crop")]

nass_data_new <- saveRDS(nass_data_new, "Shared/Data/ProcessedData/nass_data_new.rds")
```