### 2023-08-30 Codes/0_2_quickstat.Rmd
minor updates

---
title: "Get crop yield and acreage data: Corn and Soybeans"
---

# Preparation
Read the counties data

```{r }
base_counties <- readRDS(here("Shared/Data/ProcessedData/base_counties.rds"))
```

Get the list of states

```{r }
all_states <-
  base_counties$state_name %>%
  unique() %>%
  toupper()
```

Define a function to get QuickStat data

```{r }
get_quickstat_data <- function(year, data_item) {
  desc <-
    tidyUSDA::allDataItem %>%
    grep(pattern = data_item, ., value = TRUE)

  temp_data <-
    tidyUSDA::getQuickstat(
      key = key_get("usda_nass_qs_api"),
      program = "SURVEY",
      data_item = desc,
      geographic_level = "COUNTY",
      state = all_states,
      year = year,
      geometry = FALSE
    ) %>%
    #--- keep only some of the variables ---#
    dplyr::select(year, county_code, state_fips_code, short_desc, Value) %>%
    rename("state_code" = "state_fips_code")

  return(temp_data)
}
```

List of items to download

```{r }
item_list <-
  c(
    "CORN, GRAIN, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "CORN, GRAIN, IRRIGATED - ACRES HARVESTED",
    "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS, NON-IRRIGATED - YIELD, MEASURED IN BU / ACRE",
    "SOYBEANS, IRRIGATED - ACRES HARVESTED",
    "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED"
  )

# tidyUSDA::allDataItem %>%
#     grep(pattern = "PLANTED", ., value = TRUE) %>%
#     grep(pattern = "SOYBEANS", ., value = TRUE) %>%
#     grep(pattern = "IRRIGATED", ., value = TRUE)

# for (i in 1:length(item_list)) {
#   desc <-
#     tidyUSDA::allDataItem %>%
#     grep(pattern = data_item, ., value = TRUE)
#   print(desc)
# }
```

List of year-item combinations

```{r }
par_list <-
  data.table::CJ(
    data_item = item_list,
    year = as.character(1985:2018)
  )
```

# Download the data

```{r download-nass-data, eval = F}
nass_data <-
  lapply(
    seq_len(nrow(par_list)),
    function(x) {
      print(x)
      data <-
        get_quickstat_data(
          year = par_list[x, year],
          data_item = par_list[x, data_item]
        ) %>%
        data.table()
      return(data)
    }
  )

# === save the data ===#
saveRDS(nass_data, here("Shared/Data/RawData/nass_data_raw.rds"))
```

# Process the nass-data

```{r process-nass-data}
(
  nass_data_wide <-
    here("Shared/Data/RawData/nass_data_raw.rds") %>%
    readRDS() %>%
    # === drop observations with county_code == 998 ===#
    .[county_code != 998, ] %>%
    # === define state-county code ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    .[, .(sc_code, year, short_desc, Value)] %>%
    .[, short_desc := case_when(
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - YIELD") ~ "corn_ir_yield",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - YIELD") ~ "corn_nir_yield",
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - ACRES HARVESTED") ~ "corn_ir_acres",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED") ~ "corn_nir_acres",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - YIELD") ~ "soy_ir_yield",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - YIELD") ~ "soy_nir_yield",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - ACRES HARVESTED") ~ "soy_ir_acres",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED") ~ "soy_nir_acres"
    )] %>%
    .[, var_type := ifelse(
      str_detect(short_desc, "acres"),
      "acres", "yield"
    )] %>%
    # === irrigated or not? ===#
    .[, ir := ifelse(
      str_detect(short_desc, "nir"),
      "nir", "ir"
    )] %>%
    # === crop type distinction ===#
    .[, crop := ifelse(
      str_detect(short_desc, "corn"),
      "corn", "soy"
    )] %>%
    .[, short_desc := NULL] %>%
    # === long to wide ===#
    dcast(sc_code + year + crop + ir ~ var_type, value.var = "Value")
)
```

Quick Summary

```{r }
library(modelsummary)
datasummary(
  crop * ir * (Mean + SD)
  ~ acres + yield,
  data = nass_data_wide
)
```

# Save the data

```{r eval = F}
saveRDS(nass_data_wide, here("Shared/Data/ProcessedData/nass_data.rds"))
```