### 2023-08-30 Codes/2_figures_tables.rmd

---
title: "Create Figures and Tables"
author: "Taro Mieno"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)
```

# Preparation

## Load the data and results

```{r}
all_results <- readRDS("Shared/Results/all_results.rds")
```

## Set themes
```{r}
theme_fig <-
  theme_bw() +
  theme(
    axis.title.x =
      element_text(
        size = 9, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
      ),
    axis.title.y =
      element_text(
        size = 9, angle = 90, hjust = .5, vjust = .9, family = "Times"
      ),
    axis.text.x =
      element_text(
        size = 8, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
      ),
    axis.text.y =
      element_text(
        size = 8, angle = 0, hjust = 1, vjust = 0, family = "Times"
      ),
    axis.ticks =
      element_line(
        linewidth = 0.3, linetype = "solid"
      ),
    axis.ticks.length = unit(.15, "cm"),
    #--- legend ---#
    legend.text =
      element_text(
        size = 9, angle = 0, hjust = 0, vjust = 0.5, family = "Times"
      ),
    legend.title =
      element_text(
        size = 9, angle = 0, hjust = 0, vjust = 0, family = "Times"
      ),
    legend.key.size = unit(0.5, "cm"),
    #--- strip (for faceting) ---#
    strip.text = element_text(size = 9, family = "Times"),
    #--- plot title ---#
    plot.title = element_text(family = "Times", face = "bold", size = 9),
    #--- margin ---#
    # plot.margin = margin(0, 0, 0, 0, "cm"),
    #--- panel ---#
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA)
  )
```

# Summary statistics

## Boxplot of water deficit by year

```{r}
data_corn <- all_results[crop == "corn", data][[1]]
data_soy <- all_results[crop == "soy", data][[1]]

wd_data <-
  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
  .[, .(wd = mean(balance)), by = .(year, crop)] %>%
  .[, ]

yield_data <-
  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
  .[, ir_or_not := fifelse(sat_cat == "dryland", "Dryland", "Irrigated")] %>%
  .[, .(yield = mean(yield), wb = mean(balance)), by = .(year, crop, ir_or_not)]

coeff_corn <- 70

g_yield_wd_corn <-
  ggplot() +
  geom_line(
    data = yield_data[crop == "corn" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield")
  ) +
  geom_point(
    data = yield_data[crop == "corn" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield"),
    size = 0.8
  ) +
  geom_line(
    data = yield_data[crop == "corn" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield")
  ) +
  geom_point(
    data = yield_data[crop == "corn" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield"),
    size = 0.8
  ) +
  geom_line(
    data = wd_data[crop == "corn"],
    aes(y = wd / coeff_corn, x = year, color = "Water Deficit"),
    linetype = 2
  ) +
  geom_point(
    data = wd_data[crop == "corn"],
    aes(y = wd / coeff_corn, x = year, color = "Water Deficit"),
    size = 0.8
  ) +
  scale_y_continuous(
    sec.axis = sec_axis(~ . * coeff_corn, name = "Water Deficit (mm)")
  ) +
  scale_x_continuous(
    name = "Year",
    breaks = 1985:2016
  ) +
  scale_color_manual(
    name = "",
    values = c(
      "Dryland Yield" = "red",
      "Irrigated Yield" = "blue",
      "Water Deficit" = "black"
    )
  ) +
  ylab("Yield (bu/acre)") +
  xlab("Year") +
  theme_fig +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90)
  ) +
  ggtitle("A. Corn")

coeff_soy <- 200
g_yield_wd_soy <-
  ggplot() +
  geom_line(
    data = yield_data[crop == "soy" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield")
  ) +
  geom_point(
    data = yield_data[crop == "soy" & ir_or_not == "Irrigated", ],
    aes(y = yield, x = year, color = "Irrigated Yield"),
    size = 0.8
  ) +
  geom_line(
    data = yield_data[crop == "soy" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield")
  ) +
  geom_point(
    data = yield_data[crop == "soy" & ir_or_not == "Dryland", ],
    aes(y = yield, x = year, color = "Dryland Yield"),
    size = 0.8
  ) +
  geom_line(
    data = wd_data[crop == "soy"],
    aes(y = wd / coeff_soy, x = year, color = "Water Deficit"),
    linetype = 2
  ) +
  geom_point(
    data = wd_data[crop == "soy"],
    aes(y = wd / coeff_soy, x = year, color = "Water Deficit"),
    size = 0.8
  ) +
  scale_y_continuous(
    sec.axis = sec_axis(~ . * coeff_soy, name = "Water Deficit (mm)")
  ) +
  scale_x_continuous(
    name = "Year",
    breaks = 1985:2016
  ) +
  scale_color_manual(
    name = "",
    values = c(
      "Dryland Yield" = "red",
      "Irrigated Yield" = "blue",
      "Water Deficit" = "black"
    )
  ) +
  ylab("Yield (bu/acre)") +
  xlab("Year") +
  theme_fig +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90)
  ) +
  ggtitle("B. Soybean")

library(patchwork)

g_y_wd <- g_yield_wd_corn / g_yield_wd_soy

ggsave(
  file = "GitControlled/Writing/Figures/g_y_wd_all.pdf",
  g_y_wd,
  height = 6,
  width = 6
)
```

## Saturated thickness

```{r, eval = FALSE}
g_sat_dist <-
  rbind(data_corn[year <= 2016, ], data_soy[year <= 2016, ]) %>%
  .[sat != 0, ] %>%
  ggplot(data = .) +
  geom_histogram(aes(x = sat), fill = "white", color = "blue") +
  facet_grid(type ~ ., scale = "free_y") +
  ylab("Count") +
  xlab("Aquifer Thickness (meter)") +
  theme_fig
```

## Map

```{r echo = F, eval = F}
# === hpa ===#
hpa_simplified <-
  # here("../Shared/Data/ProcessedData/hp_simplified.shp") %>%
  here("Shared/Data/ProcessedData/hp_simplified.shp") %>%
  readRDS()

# === all counties ===#
data(fips_codes, package = "tidycensus")

all_counties <-
  tigris::counties() %>%
  rename(
    state_code = STATEFP,
    county_code = COUNTYFP,
    county_name = NAME
  ) %>%
  left_join(., fips_codes, by = c("state_code", "county_code")) %>%
  dplyr::select(state_code, state, county_code, county_name) %>%
  mutate(sc_code = paste0(state_code, county_code)) %>%
  data.table()

all_states <-
  tigris::states() %>%
  rename(state_code = STATEFP) %>%
  left_join(
    .,
    unique(data.table(fips_codes), by = "state"),
    by = "state_code"
  )

states_of_interest <-
  dplyr::filter(all_states, state %in% data_corn[, unique(state)])

#--- corn ---#
counties_corn_all <-
  all_counties[sc_code %in% unique(data_corn$sc_code), ] %>%
  mutate(type = "Corn") %>%
  st_as_sf()

counties_corn_ir <-
  data_corn %>%
  .[sat_cat != "dryland", ] %>%
  .[, .(sat, sc_code)] %>%
  .[, .(sat = mean(sat, na.rm = TRUE)), by = sc_code] %>%
  left_join(all_counties, by = "sc_code") %>%
  .[, type := "Corn"] %>%
  st_as_sf()

#--- soy ---#
counties_soy_all <-
  all_counties[sc_code %in% unique(data_soy$sc_code), ] %>%
  mutate(type = "Soybean") %>%
  st_as_sf()

counties_soy_ir <-
  data_soy %>%
  .[sat_cat != "dryland", ] %>%
  .[, .(sat, sc_code)] %>%
  .[, .(sat = mean(sat, na.rm = TRUE)), by = sc_code] %>%
  left_join(all_counties, by = "sc_code") %>%
  .[, type := "Soybean"] %>%
  st_as_sf()

counties_both_all <- rbind(counties_corn_all, counties_soy_all)
counties_both_ir <-
  rbind(counties_corn_ir, counties_soy_ir) %>%
  mutate(sat_meter = sat)

g_corn <-
  ggplot() +
  geom_sf(data = counties_corn_all, fill = "gray") +
  geom_sf(data = counties_corn_ir, aes(fill = sat)) +
  scale_fill_viridis_c(
    name = "Aquifer Thickness (meter)",
    guide = guide_colorbar(
      title.position = "top"
    )
  ) +
  geom_sf(data = states_of_interest, fill = NA, color = "blue") +
  geom_sf(data = hpa_simplified, fill = NA, color = "red") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm"),
    plot.title = element_text(size = 9, family = "Times"),
    legend.title = element_text(family = "Times"),
    legend.text = element_text(family = "Times")
  ) +
  ggtitle("A. Corn")

g_soy <-
  ggplot() +
  geom_sf(data = counties_soy_all, fill = "gray") +
  geom_sf(data = counties_soy_ir, aes(fill = sat)) +
  scale_fill_viridis_c(
    name = "Aquifer Thickness (meter)",
    guide = guide_colorbar(
      title.position = "top"
    )
  ) +
  geom_sf(data = states_of_interest, fill = NA, color = "blue") +
  geom_sf(data = hpa_simplified, fill = NA, color = "red") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1, "cm"),
    plot.title = element_text(size = 9, family = "Times"),
    legend.title = element_text(family = "Times"),
    legend.text = element_text(family = "Times")
  ) +
  ggtitle("B. Soybean")

g_map <- g_corn | g_soy

ggsave(file = "GitControlled/Writing/Figures/g_map.png", g_map, height = 7, width = 6, dpi = 600)
ggsave(file = "GitControlled/Writing/Figures/g_map.pdf", g_map, height = 7, width = 6)
```

# Yield response to water deficit by saturated thickness level (intensive margin)

```{r}
# all_results <-  data.table(all_results)

plot_yield_data <-
  all_results[, .(crop, yield_pred_data)] %>%
  rowwise() %>%
  dplyr::mutate(plot_data = list(
    yield_pred_data[, crop_type := crop] %>%
      .[, crop_type := case_when(
        crop_type == "corn" ~ "A. Corn",
        crop_type == "soy" ~ "B. Sybean"
      )]
  )) %>%
  dplyr::mutate(sat_cat_text_ls = list(
    plot_data[, .(sat_rank, sat_cat_text)] %>%
      .[!is.na(sat_rank), ] %>%
      unique(by = "sat_rank") %>%
      .[order(sat_rank), sat_cat_text]
  )) %>%
  dplyr::mutate(plot_data = list(
    plot_data %>%
      .[is.na(sat_cat_text), sat_cat_text := "Rainfed"] %>%
      .[, sat_cat_text := factor(sat_cat_text, levels = c("Rainfed", sat_cat_text_ls))]
  ))

g_ir_yield_both <-
  plot_yield_data %>%
  dplyr::mutate(g_yield = list(
    ggplot(plot_data) +
      geom_line(aes(
        y = y_hat,
        x = balance,
        color = factor(sat_cat_text)
      )) +
      scale_x_continuous(
        name = "Water Deficit (mm)",
        breaks = (-2:11) * 100
      ) +
      scale_y_continuous(
        name = "Yield (tonne/ha)",
        breaks = 0:(ceiling(max(plot_data$y_hat)) + 0.5),
        limits = c(NA, (ceiling(max(plot_data$y_hat)) + 0.5))
      ) +
      scale_color_discrete(
        name = "Aquifer Thickness",
        guide = guide_legend(
          title.position = "top",
          nrow = 2
        )
      ) +
      scale_fill_discrete(guide = "none") +
      theme_fig +
      coord_cartesian(xlim = c(-200, 1150)) +
      theme(
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_blank()
      )
  ))

g_ir_yield_corn <- g_ir_yield_both$g_yield[[1]] + theme(legend.position = "none")
g_ir_yield_soy <- g_ir_yield_both$g_yield[[2]]

g_wds <-
  all_results[, .(crop, data)] %>%
  dplyr::mutate(g_fig = purrr::map(
    data,
    function(data) {
      ggplot(data) +
        geom_histogram(aes(x = balance), bins = 50, col = "darkgreen", fill = NA) +
        theme_fig +
        xlab("") +
        ylab("") +
        # scale_x_continuous(breaks = (-2:11)*100) +
        coord_cartesian(xlim = c(-200, 1150))
        # +
        # # xlim(NA, max_balance) +
        # theme_void()
    }
  ))

g_wd_corn <-
  g_wds$g_fig[[1]] +
  ggtitle("A. Corn") +
  theme(
    plot.title = element_text(size = 9, family = "Times", face = "bold"),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(0, 0, 0, 0, "cm")
  )

g_wd_soy <-
  g_wds$g_fig[[2]] +
  ggtitle("B. Soybean") +
  theme(
    plot.title = element_text(size = 9, family = "Times", face = "bold"),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(0, 0, 0, 0, "cm")
    )

g_yield_response <-
  cowplot::plot_grid(
    g_wd_corn, g_ir_yield_corn, g_wd_soy, g_ir_yield_soy,
    ncol = 1,
    align = "v",
    rel_heights = c(0.15 / 2, 0.72 / 2, 0.15 / 2, 0.98 / 2)
  )

# g_yield_response <-
#   cowplot::plot_grid(
#     g_wd_corn, g_ir_yield_corn,
#     ncol = 1,
#     align = "v",
#     rel_heights = c(0.15, 0.72)
#   )

# g_yield_response_soy <-
#   cowplot::plot_grid(
#     g_wd_soy, g_ir_yield_soy,
#     ncol = 1,
#     align = "v",
#     rel_heights = c(0.15, 0.72)
#   )

cowplot::ggsave2("GitControlled/Writing/Figures/g_yield_intensive.pdf", g_yield_response, width = 6, height = 6)

#++++++++++++++++++++++++++++++++++++
#+ With confidence interval (seprately)
#++++++++++++++++++++++++++++++++++++
g_yield_with_conf <-
  plot_yield_data %>%
  dplyr::mutate(plot_data = list(
    plot_data[, `:=`(
      lower = y_hat - 1.96 * y_hat_se_modeled,
      upper = y_hat + 1.96 * y_hat_se_modeled
    )]
  )) %>%
  dplyr::mutate(g_fig = list(
      ggplot(plot_data) +
        geom_ribbon(
          aes(
            ymin = lower,
            ymax = upper,
            x = balance
          ),
          alpha = 0.3
        ) +
        geom_line(aes(
          y = y_hat,
          x = balance
        )) +
        facet_wrap(~sat_cat_text, ncol = 2) +
        scale_x_continuous(
          name = "Water Deficit (mm)",
          breaks = (-1:5) * 200
        ) +
        scale_y_continuous(
          name = "Yield (tonne/ha)",
          breaks = 0:14
        ) +
        theme_fig +
        theme(
          legend.position = "bottom",
          strip.background = element_blank()
        )
  )) %>%
  dplyr::select(crop, g_fig) %>%
  data.table()

ggsave("GitControlled/Writing/Figures/g_yield_with_conf_corn.pdf", g_yield_with_conf[crop == "corn", g_fig][[1]], width = 6, height = 3.5)
ggsave("GitControlled/Writing/Figures/g_yield_with_conf_soy.pdf", g_yield_with_conf[crop == "soy", g_fig][[1]], width = 6, height = 3.5)
```

# The impact of saturated thickness on the share of irrigated production

```{r}
plot_data_ir_share <-
  all_results[, .(crop, share_sum)] %>%
  unnest(cols = share_sum) %>%
  data.table() %>%
  .[, crop_type := case_when(
    crop == "corn" ~ "A. Corn",
    crop == "soy" ~ "B. Soybean"
  )]

g_share <-
  ggplot(plot_data_ir_share[sat <= 150, ]) +
  geom_line(aes(y = ir_share_hat, x = sat)) +
  geom_ribbon(
    aes(
      ymin = lower,
      ymax = upper,
      x = sat
    ),
    alpha = 0.4
  ) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(10, 160, by = 10)) +
  xlab("Aquifer Thickness (meter)") +
  ylab("Decline in Estimated Share of Irrigated Acres") +
  facet_wrap(crop_type ~ ., ncol = 1) +
  theme_fig +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = )
  )

ggsave(
  file = "GitControlled/Writing/Figures/g_share.pdf",
  g_share,
  height = 4,
  width = 5
)
```

# Difference in the share of irrigated production (from 130)

```{r}
g_share_dif <-
  ggplot(plot_data_ir_share[sat < 105, ]) +
  geom_line(aes(y = dif_ir_share_hat, x = sat)) +
  geom_ribbon(
    aes(
      ymin = dif_lower,
      ymax = dif_upper,
      x = sat
    ),
    alpha = 0.4
  ) +
  geom_hline(yintercept = 0, color = "red") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(10, 160, by = 10)) +
  xlab("Aquifer Thickness (meter)") +
  ylab("Estimated Share of Irrigated Acres") +
  facet_wrap(crop_type ~ ., ncol = 1) +
  theme_fig +
  theme(
    legend.position = "bottom",
    strip.background = element_blank(),
    strip.text = element_text(hjust = 0, face = "bold", family = "Times", size = )
  )

ggsave(
  file = "GitControlled/Writing/Figures/g_share_dif.pdf",
  g_share_dif,
  height = 4,
  width = 5
)
```

# Total impact of a decline in saturated thickness on average yield

```{r}
g_total_both <-
  all_results[, .(crop, avg_yield_sum)] %>%
  unnest() %>%
  data.table() %>%
  .[, crop_type := case_when(
    crop == "corn" ~ "A. Corn",
    crop == "soy" ~ "B. Soybean"
  )] %>%
  nest_by(crop_type) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(data, \(data) {
    ggplot(data) +
      geom_line(aes(y = avg_yield, x = balance, color = factor(sat))) +
      theme_fig +
      ylim(min(data$avg_yield) - 0.5, max(data$avg_yield) + 0.5) +
      ylab("Yield (tonne/ha)") +
      scale_color_discrete(
        "Aquifer Thickness (meter)",
        guide = guide_legend(
          title.position = "top",
          nrow = 1
        )
      ) +
      scale_x_continuous(name = "Water Deficit (mm)", breaks = (-2:11) * 100, limits = c(-200, 1200)) +
      theme(legend.position = "bottom")
  }))

g_total_corn <-
  g_total_both$g_fig[[1]] +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  ylim(6, 13) +
  ggtitle("A. Corn")

g_total_soy <-
  g_total_both$g_fig[[2]] +
  ylim(1.5, 4) +
  ggtitle("B. Soybean")

g_total <- g_total_corn / g_total_soy

ggsave(
  file = "GitControlled/Writing/Figures/g_total_impact.pdf",
  g_total,
  height = 6,
  width = 6
)
```

# Test the significance of the difference in irrigated yield between the aquifer thickness categories

```{r}
g_dif <-
  all_results[, .(crop, yield_dif_test, sat_text_data)] %>%
  rowwise() %>%
  dplyr::mutate(plot_data = list(
    yield_dif_test[, crop_type := crop] %>%
      .[, crop_type := case_when(
        crop_type == "corn" ~ "A. Corn",
        crop_type == "soy" ~ "B. Sybean"
      )] %>%
      setnames("comp_q", "sat_rank") %>%
      sat_text_data[., on = "sat_rank"] %>%
      .[, dif_upper := dif_hat + 1.96 * se] %>%
      .[, dif_lower := dif_hat - 1.96 * se]
  )) %>%
  dplyr::mutate(sat_cat_text_ls = list(
    plot_data[, .(sat_rank, sat_cat_text)] %>%
      .[!is.na(sat_rank), ] %>%
      unique(by = "sat_rank") %>%
      .[order(sat_rank), sat_cat_text]
  )) %>%
  dplyr::mutate(plot_data = list(
     plot_data[, sat_cat_text := factor(sat_cat_text, levels = sat_cat_text_ls)]
  )) %>%
  dplyr::mutate(g_fig = list(
    ggplot(plot_data) +
      facet_grid(. ~ sat_cat_text) +
      geom_ribbon(
        aes(
          x = balance,
          ymin = dif_lower,
          ymax = dif_upper
        ),
        fill = "blue",
        alpha = 0.3
      ) +
      geom_line(
        aes(
          x = balance,
          y = dif_hat
        ),
        color = "red"
      ) +
      geom_hline(yintercept = 0, col = "black") +
      scale_color_discrete(name = "") +
      scale_x_continuous(breaks = (-1:6) * 200) +
      ylab("Difference in Yield (tonne/ha)") +
      xlab("Water Deficit (mm)") +
      theme_fig +
      theme(
        strip.background = element_blank(),
        legend.position = "bottom",
        panel.spacing = unit(1, "lines")
      )
  ))

g_ir_yield_dif <-
  cowplot::plot_grid(
    g_dif$g_fig[[1]] + ggtitle("A. Corn") + ylim(-3, 3),
    g_dif$g_fig[[2]] + ggtitle("B. Soybean") + ylim(-3, 3),
    ncol = 1,
    align = "v",
    rel_heights = c(0.5, 0.5)
  )

cowplot::ggsave2("GitControlled/Writing/Figures/g_ir_yield_dif.pdf", g_ir_yield_dif, width = 6, height = 5)
```

# Testing of the significance of yield difference (average total impact)

```{r}
g_avg_yield_dif <-
  all_results[, .(crop, avg_yield_sum)] %>%
  unnest(cols = c(avg_yield_sum)) %>%
  data.table() %>%
  .[, sat_text := paste0(
    "Aquifer Thickness = ",
    format(sat, digits = 3),
    "m"
  )] %>%
  .[sat != min(sat), ] %>%
  nest_by(crop) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(data, \(data){
    ggplot(data) +
      geom_ribbon(aes(x = balance, ymin = dif_lower, ymax = dif_upper), alpha = 0.3, fill = "blue") +
      geom_line(aes(x = balance, y = dif_avg_yield), color = "red") +
      geom_hline(yintercept = 0) +
      facet_grid(. ~ sat_text, scales = "free_y") +
      xlab("Water Deficit (mm)") +
      ylab("Difference in Yield (tonne/ha)") +
      xlim(-200, 1200) +
      theme_fig +
      theme(
        strip.background = element_blank(),
        strip.text = element_text(size = 9, family = "Times"),
        panel.spacing.y = unit(1, "cm")
      )
  }))

g_ay_corn <-
  g_avg_yield_dif$g_fig[[1]] +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(breaks = -2:5, limits = c(-2, 6)) +
  ggtitle("A. Corn")

g_ay_soy <-
  g_avg_yield_dif$g_fig[[2]] +
  scale_y_continuous(breaks = -2:2, limits = c(-2, 2)) +
  ggtitle("B. Soybean")

g_ay <- g_ay_corn / g_ay_soy

ggsave("GitControlled/Writing/Figures/g_avg_yield_dif.pdf", g_ay, width = 6, height = 5)
```

# The impact of water deficit on average yield by aquifer thickness category with confidence interval

```{r}
g_ay_response <-
  all_results[, .(crop, avg_yield_sum)] %>%
  unnest(cols = c(avg_yield_sum)) %>%
  data.table() %>%
  .[, sat_text := paste0(
    "Aquifer Thickness = ",
    format(sat, digits = 3),
    "m"
  )] %>%
  .[sat != min(sat), ] %>%
  nest_by(crop) %>%
  data.table() %>%
  dplyr::mutate(g_fig = purrr::map(data, \(data) {
    ggplot(data) +
      geom_ribbon(aes(x = balance, ymin = lower, ymax = upper), alpha = 0.3) +
      geom_line(aes(x = balance, y = avg_yield)) +
      facet_grid(. ~ sat_text, scales = "free_y") +
      ylim(min(data$lower) * 0.7, max(data$upper) + 1) +
      xlab("Water Deficit (mm)") +
      ylab("Yield (tonne/ha)") +
      theme_fig +
      theme(
        strip.background.x = element_blank(),
        panel.spacing = unit(0.1, "cm")
      )
  }))

g_ay_response_corn <-
  g_ay_response$g_fig[[1]] +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  ggtitle("A. Corn")

g_ay_response_soy <-
  g_ay_response$g_fig[[2]] +
  ggtitle("B. Soybean")

g_ay_response <- g_ay_response_corn / g_ay_response_soy

ggsave("GitControlled/Writing/Figures/g_avg_yield.pdf", g_ay_response, width = 6, height = 3.5)
```