### 2023-10-31 Codes/0_3_get_sat.Rmd

---
title: "Assign saturated thickness values"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

# Preparation
Import saturated thickness data

```{r }
sat_temp <- raster("Data/data-raw/saturated_thickness/SatThick_19851.tif")
```

HPA counties

```{r }
hpa_counties <-
  here("Data/data-processed/base_counties.rds") %>%
  readRDS() %>%
  .[in_hpa == 1, ] %>%
  st_as_sf() %>%
  st_transform(projection(sat_temp))
```

Define function to extract saturated thickness for each county

```{r }
get_sat <- function(y) {
  # y <- 2000
  sat_temp <-
    paste0("Data/data-raw/saturated_thickness/SatThick_", y, "1.tif") %>%
    raster()

  sat_mean <-
    exact_extract(sat_temp, hpa_counties, "mean")

  data_return <-
    data.table(
      sat = sat_mean,
      sc_code = hpa_counties$sc_code,
      year = y
    )

  return(data_return)
}
```

Extract saturated thickness values (parallelized)

```{r extract-sat, eval = F}
sat_all <-
  mclapply(1985:2016, get_sat, mc.cores = 12) %>%
  rbindlist() %>%
  .[sat < 0, sat := 0]
```

# Save the data

```{r save, eval = F}
saveRDS(sat_all, here("Data/data-processed/sat_all.rds"))
```