### 2023-10-31 Codes/0_6_merge_datasets.rmd

# Combine all the datasets

## Read the datasets

```{r}
base_counties <- readRDS(here("Data/data-processed/base_counties.rds"))

#' Weather
weather_data <- readRDS(here("Data/data-processed/summrized_gridMET.rds"))

#' Saturated thickness
sat_all <- readRDS(here("Data/data-processed/sat_all.rds"))

#' Yield, acreage data
nass_data <- readRDS(here("Data/data-processed/nass_data.rds"))

#* SSURGO
ssurgo_soil <- readRDS(here("Data/data-processed/ssurgo_soil.rds"))
```

## combine all

```{r}
merged_data <-
  base_counties[nass_data, on = c("sc_code")] %>%
  weather_data[., on = c("sc_code", "year")] %>%
  sat_all[., on = c("sc_code", "year")] %>%
  .[order(sc_code, year), ] %>%
  ssurgo_soil[, on = "sc_code"]
```

# Create and redefine new variables, select observations

```{r}
final_data <-
  merged_data %>%
  #--- yield unit conversion (bu/acre to ton/ha)---#
  .[, yield := yield * 12.553 / 200] %>%
  # === if not in_hpa and irrigated, drop them ===#
  # irrigated, but saturated thickness not observed. These
  # observations cannot be used.
  .[!(in_hpa == 0 & ir == "ir"), ] %>%
  # === if not irrigated, set sat = 0 ===#
  .[ir == "nir", sat := 0] %>%
  # === if sat == 0, but indicated irrigated ===#
  # add 0.1 to distinguish dryland production, where
  # sat is set at 0 above
  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
  # === drop if yield or acres are not observed ===#
  .[!is.na(yield) & !is.na(acres), ] %>%
  # === if sat is missing, drop ===#
  #' just 2017 and 2018 data because saturated thickness data
  #' is observed only until 2016
  .[!is.na(sat), ] %>%
  #--- indicator of whether dryland (rainfed) production or not ---#
  .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
  #--- sc_code and dry_or_not (used for FE regression) ---#
  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
  #--- state-year combination ---#
  .[, state_year := paste0(state_name, "_", year)] %>%
  # === number of observations by state-county code ===#
  .[, num_obs := .N, by = sc_dry] %>%
  # .[, num_obs := .N, by = sc_code] %>%
  # === keep only if more than 10 observations are available ===#
  .[num_obs >= 10, ] %>%
  # === order (not necessary) ===#
  .[order(sc_code, year), ]
```

# Save

```{r}
saveRDS(final_data, here("Data/data-processed/final_data.rds"))
```