### 2023-10-31 Codes/master.rmd
functions.R organized

# Data Preparation

```{r, eval = FALSE}
knitr::purl(here("Codes/0_1_hpa_counties.Rmd"), here("Codes/0_1_hpa_counties.R"))
knitr::purl(here("Codes/0_2_quickstat.Rmd"), here("Codes/0_2_quickstat.R"))
knitr::purl(here("Codes/0_3_get_sat.Rmd"), here("Codes/0_3_get_sat.R"))
knitr::purl(here("Codes/0_4_get_weather.Rmd"), here("Codes/0_4_get_weather.R"))
knitr::purl(here("Codes/0_5_ssurgo.Rmd"), here("Codes/0_5_ssurgo.R"))
knitr::purl(here("Codes/0_6_merge_datasets.Rmd"), here("Codes/0_6_merge_datasets.R"))
```

## Find counties in HPA
```{r, eval = FALSE}
source(here("Codes/0_1_hpa_counties.R"))
```

## Downlod yield and acreage data from USDA-NASS Quickstat

```{r}
source(here("Codes/0_2_quickstat.R"))
```

## Assign saturated thickness values to each county-year

```{r, eval = FALSE}
source(here("Codes/0_3_get_sat.R"))
```

## Download gridMET data and summarize by county-year

+ Downloading all the gridMET data and processing them take a while.
+ The total size of the gridMET files is about 35GB.

```{r, eval = FALSE}
source(here("Codes/0_4_get_weather.R"))
```

## Download SSURGO data and summarize by county

+ Running this takes a while.
+ SSURGO can be erratic and you may face downloading errors at times

```{r, eval = FALSE}
source(here("Codes/0_5_ssurgo.R"))
```

## Merge datasets, defined new variables, and drop observations

```{r, eval = FALSE}
source(here("Codes/0_6_merge_datasets.R"))
```

# Analysis

```{r}
knitr::purl(here("Codes/1_analysis.Rmd"), here("Codes/1_analysis.R"))
source(here("Codes/1_analysis.R"))
```

# Create figures

```{r}
knitr::purl(here("Codes/2_figures_tables.Rmd"), here("Codes/2_figures_tables.R"))
source(here("Codes/2_figures_tables.R"))
```