### 2023-11-01 Codes/0_1_hpa_counties.Rmd
manuscript.rnw created

---
title: "Prepare base-county data"
author: "Taro Mieno"
output:
  html_document:
    number_sections: yes
    theme: flatly
    highlight: zenburn
    toc_float: yes
    toc: yes
    toc_depth: 3
geometry: margin=1in
---

# Identify counties the overlaps with the HPA boundary
## Read the HPA border shapefile

```{r Read-HPA, cache = TRUE}
hpa <- st_read(here("Data/data-raw/hp_bound2010.shp"))
```

Map-of-HPA

```{r, eval = FALSE}
tm_shape(st_simplify(hpa)) +
  tm_polygons()
```

## Get all the US counties as an sf

```{r Get-US-counties}
(
  US_county <- tigris::counties(cb = TRUE) %>%
    st_as_sf() %>%
    st_transform(st_crs(hpa)) %>%
    rename("state_code" = "STATEFP", "county_code" = "COUNTYFP") %>%
    dplyr::select(state_code, county_code)
)
```

## Identify counties that overlaps with `hpa`

```{r Identify-overlap, cache = TRUE}
hpa_counties <- US_county[hpa, ] %>%
  mutate(whole_area = st_area(.) %>% as.numeric())
```

Map-of-HPA and the intersecting counties

```{r eval = FALSE}
tm_shape(st_simplify(hpa)) +
  tm_polygons() +
  tm_shape(hpa_counties) +
  tm_polygons(col = "blue", alpha = 0.4)
```

## Find the area of the portion that is intersecting with `hpa`
Intersect HPA and counties in HPA

```{r intersect-HPA-county}
intersected_portion <-
  st_intersection(hpa, hpa_counties) %>%
  # === get the intersecting area ===#
  mutate(area = st_area(.) %>% as.numeric()) %>%
  data.table() %>%
  .[, .(interesected_area = sum(area)), by = .(state_code, county_code)]
```

Merge the intersection data

```{r }
hpa_counties <-
  left_join(
    hpa_counties,
    intersected_portion,
    by = c("state_code", "county_code")
  ) %>%
  # === find the proportion of the county area that overlaps with HPA ===#
  mutate(ir_area_ratio = interesected_area / whole_area) %>%
  data.table() %>%
  .[, geometry := NULL]
```

Histogram of coverage ratio

```{r eval = FALSE}
hpa_counties$ir_area_ratio %>% hist()
```

Get the list of the states

```{r }
hpa_states <- hpa_counties$state_code %>% unique()
```

# Additional counties 
Additional state codes:
these supplement the dryland production data for understanding the impact of drought severity
on dryland crop production

```{r }
data(fips_codes)

additional_states <-
  fips_codes %>%
  data.table() %>%
  unique(by = "state_name") %>%
  .[state_name %in% c("Illinois", "Iowa", "Missouri", "Indiana"), ] %>%
  .[, state_code]

all_states <- c(hpa_states, additional_states)
```

# All counties
**Get the state abbreviation names**

```{r }
fips_codes <-
  fips_codes %>%
  data.table() %>%
  .[, .(state, state_code, state_name, county_code)]
```

Get all the counties in the target states

```{r }
(
  all_counties <-
    dplyr::filter(US_county, state_code %in% all_states) %>%
    data.table() %>%
    # === merge with the data on counties in HPA ===#
    hpa_counties[., on = c("state_code", "county_code")] %>%
    # === merge with FIPS codes data ===#
    fips_codes[., on = c("state_code", "county_code")] %>%
    # === indicator of whether intersecting with HPA or not ===#
    .[, in_hpa := 0] %>%
    .[!is.na(ir_area_ratio), in_hpa := 1] %>%
    # === generate state-code variable ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    # === keep only the relevant variables ===#
    .[, .(state, state_name, sc_code, ir_area_ratio, geometry, in_hpa)]
)
```

Map of all the counties: 
Note: not all the counties are used in our eventual analysis

```{r map-all-counties, eval = FALSE}
tm_shape(st_as_sf(all_counties)) +
  tm_polygons() +
  tm_shape(
    tigris::states() %>%
      filter(STATEFP %in% all_states)
  ) +
  tm_borders(col = "red")
```

# save the county data

```{r }
saveRDS(all_counties, here("Data/data-processed/base_counties.rds"))
```