### 2023-11-03 Data/metadata.Rmd
minimum set of datasets added

---
title: Metadata
author: Taro Mieno
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    number_sections: true
---

# Data/data-raw

+ **hp_bound2010.shp**: HPA aquifer boundary with no attribute (shape file)
+ **metadata_elevationdata.nc**: ~ 4km elevation data from the gridMET website (netcdf file)
+ **nass_data_raw.rds** (`data.table`): downloaded from USDA NASS using the R `tidyusda` package (generated in **Codes/0_2_quickstat.Rmd**)
  + `year`: year
  + `county_code`: county fips code
  + `state_code`: state fips code
  + `short_desc`: description of what `Value` variable represents
  + `Value`: value

## data-raw -> girdMET

Files in this folder stores the raw gridMET data saved as netcdf files. They follow the following naming convention:

"weather_variable"_"year.nc"

Single files records daily values for the corresponding variable.

Here are the list of the variables:

+ "pr": precipitation
+ "rmin": min relative humidity (%)
+ "rmax": max relative humidity (%)
+ "srad": radiation (W/M^2)
+ "tmmn": tmin (Kelvin)
+ "tmmx": tmax (Kelvin)
+ "vs": wind speed,
+ "pet": evapotranspiration

See **Codes/04_get_weather.Rmd** for the codes to download them.

## data-raw -> saturated_thickness

Files in this folder stores the aquifer thickness data saved as geotif files. They follow the following naming convention:

"SatThick"_"year.tif"

# Data/data-processed

+ **base_counties.rds** (`data.table`): boundary and attributes of counties (generated in **0_1_hpa_counties.Rmd**)
  - `state`: state (abbreviated)
  - `state_name`: state (full)
  - `sc_code`: State and County FIPS codes combined into one
  - `ir_area_ratio`: the ratio of the county area that is overlapping with the HPA
  - `geometry`: simple feature column reprsenting county boundary
  - `in_hpa`: whether intersecting with the HPA or not (1 if intersected, 0 otherwise)
+ **gridMET.rds** (`data.table`): daily weather data from gridMET (generated in **0_4_get_weather.Rmd**)
  + `date`: date 
  + `tmin`: minimum temperature
  + `tmax`: maximum temperature
  + `prcp`: precipitation
  + `et0`: reference evapotranspiration
  + `sc_code`: State and County FIPS codes combined into one
+ **nass_data.rds** (`data.table`): **nass_data_raw.rds** reshaped to a long format (generated in **0_2_quickstat.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
  + `crop`: crop type (corn or soybean)
  + `ir`: irrigated or rainfed ("ir" if irrigated, "nir" if not.)
  + `acres`: harvested acres
  + `yield`: crop yield (bu/acre)
+ **sat_all.rds** (`data.table`): aquifer (saturated) thickness  (generated in **0_3_get_sat.Rmd**)
  + `sat`: aquifer (saturated) thickness level
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
+ **ssurgo_soil.rds** (`data.table`): SSURGO soil data combined into one (generated in **0_5_ssurgo.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `sandtotal_r`: sand percentage
  + `silttotal_r`: silt percentage
  + `awc_r`: water holding capacity
+ **summarized_gridMET.rds** (`data.table`): gridMET data summerized by production season (generated in **0_4_get_weather.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
  + `balance`: water balance (deficit) = `et0` - `prcp`
  + `et0`: total reference evapotranspiration (mm) during the production season defined as May through September
  + `prcp`: total precipitation (mm) during the production season defined as May through September
+ **final_data.rds** (`data.table`): dataset for analysis (generated in **0_6_merge_datasets.Rmd**)
  + `sc_code`: State and County FIPS codes combined into one
  + `year`: year
  + `sat`: aquifer (saturated) thickness (m)
  + `balance`: water balance (defict) = `et0` - `prcp`
  + `et0`: total reference evapotranspiration (mm) during the production season defined as May through September
  + `prcp`: total precipitation (mm) during the production season defined as May through September
  + `state`: abbreviated state name
  + `state_name`: full state name
  + `ir_area_ratio`: the ratio of the county area that is overlapping with the HPA
  + `geometry`: simple feature column reprsenting county boundary
  + `in_hpa`: whether intersecting with the HPA or not (1 if intersected, 0 otherwise)
  + `ir`: irrigated or rainfed ("ir" if irrigated, "nir" if not.)
  + `acres`: harvested acres
  + `yield`: crop yield (ton/ha)
  + `sandtotal_r`: sand percentage
  + `silttotal_r`: silt percentage
  + `awc_r`: water holding capacity
  + `sc_dry`: `sc_code`_`ir` (concatenation of `sc_code` and `ir`)
  + `state_year`: `state`_`year` (concatenation of `state` and `year`)
+ **hpa_simplified.rds** (`sf`): boundary of HPA simplified from **hp_bound2010.shp**, used only for the aquifer thickness map (generated in **2_figures_tables.Rmd**)
+ **states.rds** (`sf`): state boundary, used only for the aquifer thickness map (generated in **2_figures_tables.Rmd**)
+ **counties.rds** (`sf): county boundary, used only for the aquifer thickness map (generated in **2_figures_tables.Rmd**)

```{r, eval = FALSE, echo = FALSE}
readRDS("Data/data-processed/final_data.rds")
```

## Data/data-processed/SSURGO

This folder houses a series of SSURGO soil data files. Each file pertains to a single county and contains the same variables as found in **ssurgo_soil.rds**:

+ `sc_code`: State and County FIPS codes combined into one
+ `sandtotal_r`: sand percentage
+ `silttotal_r`: silt percentage
+ `awc_r`: water holding capacity

Indeed, `ssurgo_soil.rds` is an aggregation of the files contained in this folder.

Files for individual counties are saved separately due to the instability of the SSURGO website.