### 2023-11-04 Codes/3_split_manuscript.Rmd
a code to split the all-in-one manuscript to the main and SI parts.

```{r}
master <- readLines(here::here("Writing/nw_manuscript.Rnw"))
```

# Main

```{r}
main_end_line <- which(str_detect(master, "%= Supplementary Information")) - 2

which(str_detect(main, "pathway-reg"))

main <-
  master[1:main_end_line] %>%
  #--- this assumes the section ordering is correct and do not change any more ---#
  gsub("\\\\ref\\{pathway-reg\\}", "A", .) %>%
  gsub("\\\\ref\\{reg-conf\\}", "B", .) %>%
  gsub("\\\\ref\\{test-dif-share\\}", "C", .) %>%
  gsub("\\\\ref\\{spline-basis\\}", "D", .) %>%
  gsub("\\\\ref\\{county-fe\\}", "E", .)

writeLines(main, here::here("Writing/nw_main.Rnw"))
```

# SI