### 2023-11-07 Codes/3_gen_manuscript.Rmd
manuscript split into two: main and SI

```{r}
master <- readLines(here::here("Writing/nw_manuscript.Rnw"))
```

# Main

```{r}
main_end_line <- which(str_detect(master, "%= Supplementary Information")) - 2

end_doc <- which(str_detect(master, "end\\{document\\}"))

main <-
  master[c(1:main_end_line, end_doc)] %>%
  #--- this assumes the section ordering is correct and do not change any more ---#
  gsub("\\\\ref\\{pathway-reg\\}", "A", .) %>%
  gsub("\\\\ref\\{reg-conf\\}", "B", .) %>%
  gsub("\\\\ref\\{test-dif-share\\}", "C", .) %>%
  gsub("\\\\ref\\{spline-basis\\}", "D", .) %>%
  gsub("\\\\ref\\{county-fe\\}", "E", .) %>%
  gsub("\\\\ref\\{fig:irrigated-yield-ind-corn\\}", "S.1", .) %>%
  gsub("\\\\ref\\{fig:irrigated-yield-ind-soy\\}", "S.2", .)

writeLines(main, here::here("Writing/nw_main.Rnw"))
```

# SI

```{r}
main_start_line <- which(str_detect(master, "maketitle"))

main_end_line <- which(str_detect(master, "appendix-appendix"))

bib_line <- which(str_detect(master, "DRA.bib"))

si <-
  master[-c(main_start_line:main_end_line)] %>%
  gsub("\\\\ref\\{main\\}", "Main", .)

si_end_doc <- which(str_detect(si, "end\\{document\\}"))

si <- c(si[-si_end_doc], master[bib_line], si[si_end_doc])

writeLines(si, here::here("Writing/nw_si.Rnw"))
```